---
title: "Hypothesis regressions for FINDONOR- CHANGES IN SELF-RATED HEALTH DURING THE STUDY AND IT'S RELATION TO IRON BIOMARKERS "
author: "Mikko Arvas"
date: "`r Sys.time()`"
output: html_document
---

TODO:
-jpg plotting

```{r setup, include=FALSE}
#echo "rmarkdown::render('hypothesis_regressions.Rmd', clean=TRUE,output_format='html_document',output_file='../results/hypothesis_regressions.html')" | R --slave
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
library(tidyverse)
library(epitools)
library(ez)
library(lubridate)
library(ggbeeswarm)
library(tidymodels)
library(broom)
library(GGally)
library(ggfortify)
library(car)
library(boot)
library(tableone)
library(kableExtra)
```

# Executive summary

This file includes the code for computing the regression results regarding the predictors of the evolution of self-related health during the FinDonor study for the 

# Load the data

We load the data and assign groups as they were assigned in the first part of the study. 

```{r loading}
#Load questionnaire data
load(file = "../data/final_data.rdata") 
final_data <- data

#load measurement data
load(file = "../data/quest_2_donor_data.rdata")

FD_activity_biomarker_data <- data %>% 
  rename(Ferritin_beginning = Ferritin)

first_questionnaire_data <- final_data %>% filter(questionnaire == "First")
second_questionnaire_data <- final_data %>% filter(questionnaire == "Second")
rm(data)
```

# Preprocessing



We remove donors with very high ferritin (/Ferritin > 400) and high CRP (> 30)

We compute for each donor their wight at first donation and then the difference in reported BMI between the first questionnaire and the last questionnaire. 

We then build a categorical variable for smoking status. This variable will need to be dummy coded into 3 variables if it is used as a regression variable.

* Never smoker: non smoker at both time-points
* Always smoker: smoker at both time-points
* New smoker: non-smoker at first but smoker at second
* Former smoker: smoker at first but non-smoker at second.

We will apply this coding as it was applied in a recent [epidemiological study](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6290273/) in the Japanese population. Importantly we will consider only changes from smoker to non-smoker, and not changes fron "daily" to "occasional" smoker, as was done in the previously cited paper(NOTE: they had much more precise quantification of smoking behavior (see [questionnaire](https://epi.ncc.go.jp/files/00_common/questionnaire/english/BL_Cohort_I_questionnaire_English.pdf) ) and a much larger N and still used a binary smoking/non-smoking catergorisation). However, in the modelling stage we find that smoking can not be used in the models.

We also add age to the dataframe.

The data object FD_activity_biomarker_data includes all necessary variables for implenting  the model.

```{r}



FD_activity_biomarker_data <-
  final_data %>% 
    group_by(donor) %>%
  # Get staring weight
    filter(questionnaire =="First") %>% 
    rename(starting_weight = weight) %>% 
    dplyr::select(donor, starting_weight) %>% 
    full_join(final_data, by = "donor") %>% 
  # Get difference in BMI
    dplyr::select(donor, starting_weight, BMI, questionnaire) %>% 
    spread(key = questionnaire, value = BMI ) %>% 
    mutate(BMI_diff = Second - First) %>% 
    dplyr::select(-First, -Second) %>% 
    full_join(final_data, by = "donor") %>% 
  # GEt smoking status evolution
    dplyr::select(donor, starting_weight, BMI_diff, smoking, questionnaire) %>% 
    spread(key = questionnaire, value = smoking) %>% 
    mutate(smoking_behavior = case_when(First == "no" & Second == "no" ~"never smoker",
                                        First == "no" &  Second %in% c("daily", "sometimes") ~ "new smoker",
                                        First %in% c("daily", "sometimes") & Second %in% c("daily", "sometimes") ~ "always smoker",
                                        First %in% c("daily", "sometimes") & Second == "no" ~"Former smoker",
                                        TRUE ~"NA")) %>%  
    dplyr::select(-First, -Second) %>% 
    full_join(final_data, by = "donor") %>%
    dplyr::select(donor, starting_weight, BMI_diff, health, smoking_behavior, questionnaire) %>% 
  # Get health status evolution
     spread(key = questionnaire, value = health) %>% 
     mutate(health_evolution = case_when(First < Second ~ "Improved",
                                  First == Second ~ "Stable",
                                  Second < First ~ "Worsened",
                                  TRUE ~"NA")) %>% 
    dplyr::select(-First, -Second) %>% 
  # Add number of months between donation events  
    full_join(final_data, by = "donor") %>%
    dplyr::select(donor, starting_weight, BMI_diff, health_evolution, smoking_behavior, date, questionnaire,group) %>% 
    spread(key = questionnaire, value = date) %>% 
  mutate(interval_between_quest = interval(First, Second) %>% 
                                    as.duration() %>% 
           as.numeric( "months") %>% 
           round(0)) %>% 
    dplyr::select(-First, -Second) %>% 
  # Add donor age
    left_join(first_questionnaire_data %>%
              dplyr::select(donor, age),
            by = "donor") %>%
      filter( group != "Women_pre_menop_no_mens") %>% 
  # Exckude donors with no measure of health evolution
  left_join(FD_activity_biomarker_data, by = "donor") %>%
  filter(health_evolution != "NA") %>% 
  drop_na(starting_weight,BMI_diff ) %>% 
  filter(Ferritin_beginning <= 400 & CRP < 30)

final_data <-
  final_data %>% 
    filter(donor %in% FD_activity_biomarker_data$donor)
```


We first count how many donors we have in each group. 

```{r}
final_data %>% 
  filter(questionnaire == "First") %>% 
   count(group)
```


# Regressions

## Regression data

We select the variabels that will be included in the models.

We also run some trasnformations to make the OR more interpretable:

* Ferritin is transformed to $$log(Ferritin)/log(2)$$ so that a unit increase in the transformed variable corresponds to a doubling of ferritin
* weight, age and Hb were divided by 5 so that a unit increase in the transformed varaible corresponds to an increase in 5 kg or 5 years or 5 Hb points. 

```{r}


regression_data <-
  FD_activity_biomarker_data %>% 
  inner_join(first_questionnaire_data %>% 
               dplyr::select(donor, health),
             by = "donor") %>% 
  mutate(health_very_good = ifelse(health == "Very_good",1,0),
         health_excellent = ifelse(health == "Excellent", 1, 0),
         health_poor = ifelse(health == "Poor", 1, 0),
         health_satisfactory = ifelse(health == "Satisfactory", 1, 0)
         ) %>% 
  mutate(health_outcome = ifelse(health_evolution == "Worsened", 1,0)) %>%
  mutate(health_outcome_neg = ifelse(health_evolution == "Improved", 1,0)) %>%
  mutate(donation_frequency = nb_donations_between_questionnaires/(interval_between_quest/12) ) %>% 
  mutate(log_Ferritin_beginning = log(Ferritin_beginning)/log(2),
         starting_weight_o =starting_weight, 
         starting_weight = starting_weight/5,
         age_o = age,
         age = age/5,
         Hb_beginning_o = Hb_v,
         Hb_beginning = Hb_v/5) %>% 
  dplyr::select(donor, starting_weight,BMI_diff, age, group, log_Ferritin_beginning, Hb_beginning,
                donation_frequency,health_poor,health_satisfactory, health_very_good,health_excellent,health_outcome,health_outcome_neg, smoking_behavior,Ferritin_beginning,age_o,starting_weight_o,Hb_beginning_o )

  
regression_data$smoking_behavior <- factor(regression_data$smoking_behavior)
regression_data <- droplevels(regression_data)
file <- "../results/hypregdata_unscaled.rdata"
save(regression_data,file=file)


```

```{r}
regression_data %>% group_by(group) %>% 
     count()
```


# Table one

```{r}


# myVars <- c( "age_o", "starting_weight_o" ,"BMI_diff", "Hb_beginning_o", "Ferritin_beginning",
#              "donation_frequency","initial_health","final_health")

myVars <- c(   "Age" ,
  "Initial weight (kg)" ,
  "BMI difference (second - first)" ,
  "Initial hemoglobin (g/l)" ,
  "Initial ferritin (ug/l)" ,
  "Donation frequency (yearly)" ,
  "Inital health rating",
  "Final health rating")


non_normal_vars <- c("Initial ferritin (ug/l)")

health1<- first_questionnaire_data %>% dplyr::select(donor,health) %>% rename(initial_health=health)
health2<- second_questionnaire_data %>% dplyr::select(donor,health) %>% rename(final_health=health)

table1data <- regression_data %>% left_join(health1) %>% left_join(health2) 
table1data <- table1data %>% dplyr::select(group,age_o,starting_weight_o,BMI_diff,Hb_beginning_o,Ferritin_beginning,donation_frequency,initial_health,final_health) %>%
  rename(
  "Age" = age_o,
  "Initial weight (kg)" = starting_weight_o ,
  "BMI difference (second - first)" = BMI_diff,
  "Initial hemoglobin (g/l)" = Hb_beginning_o,
  "Initial ferritin (ug/l)" = Ferritin_beginning,
  "Donation frequency (yearly)" =donation_frequency,
  "Inital health rating"=initial_health,
  "Final health rating"= final_health
)


summary_table <- CreateTableOne(data = table1data,vars=myVars, strata = "group",test = FALSE)
  
tab3Mat <- print(summary_table, nonnormal = non_normal_vars,vars=myVars, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)


tab3Mat %>% 
  kable()
  write.csv(tab3Mat, file = paste0("../results/table_1_population_",gsub("-","",as.Date(Sys.time())),".csv"))
```




## Pre-menopausal women 

```{r}
regression_data_pre_menop_women <-
  regression_data %>% 
  filter(group == "Pre_menopause_women") %>% 
  dplyr::select(donor, age, starting_weight, BMI_diff,
                log_Ferritin_beginning, Hb_beginning, donation_frequency) %>% 
  gather(key = key, value = value, -donor) %>% 
  group_by(key) %>% 
  mutate(value_scale = scale(value, scale = FALSE)[,1]) %>% 
  dplyr::select(-value) %>% 
  spread(key = key, value = value_scale) %>% 
  inner_join(regression_data %>% 
               dplyr::select(donor,health_poor,health_satisfactory, health_very_good, health_excellent, health_outcome,health_outcome_neg,smoking_behavior),
             by = "donor")

```

### Correlograms

```{r echo=FALSE}

ggpairs(regression_data_pre_menop_women, 
        columns = c("health_outcome", "age", "starting_weight",  "BMI_diff","Hb_beginning", "log_Ferritin_beginning",  "donation_frequency"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)
```

```{r}
logit_pre_menop_women <- glm(health_outcome ~ age + starting_weight + BMI_diff + 
                               log_Ferritin_beginning + Hb_beginning +
                              donation_frequency + health_very_good + health_excellent,
                             data=regression_data_pre_menop_women,
                            family="binomial")
 
 summary(logit_pre_menop_women)
```

```{r}
plot(logit_pre_menop_women)
```

```{r}
vif(logit_pre_menop_women)
```

```{r}
tidy_logit_pre_menop_women <-
  logit_pre_menop_women %>% 
  tidy %>% 
  mutate(OR = exp(estimate),
         group = "Pre_menopause_women")
```

## Post-menopausal women 

```{r}
regression_data_post_menop_women <-
  regression_data %>% 
  filter(group == "Post_menopause_women") %>% 
  dplyr::select(donor, age, starting_weight, BMI_diff,
                log_Ferritin_beginning,Hb_beginning,donation_frequency) %>% 
  gather(key = key, value = value, -donor) %>% 
  group_by(key) %>% 
  mutate(value_scale = scale(value, scale = FALSE)[,1]) %>% 
  dplyr::select(-value) %>% 
  spread(key = key, value = value_scale) %>% 
  inner_join(regression_data %>% 
               dplyr::select(donor, health_poor,health_satisfactory,health_very_good, health_excellent, health_outcome,health_outcome_neg,smoking_behavior),
             by = "donor")

```

### Correlograms

```{r echo=FALSE}

ggpairs(regression_data_post_menop_women, 
        columns = c("health_outcome", "age", "starting_weight",  "BMI_diff","Hb_beginning", "log_Ferritin_beginning",  "donation_frequency"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)
```


### Models and results

```{r}
logit_post_menop_women <- glm(health_outcome ~ age + starting_weight + BMI_diff + 
                                log_Ferritin_beginning + Hb_beginning + donation_frequency +
             health_very_good + health_excellent,
             data=regression_data_post_menop_women,
             family="binomial")
 
 summary(logit_post_menop_women)
```


```{r}
tidy_logit_post_menop_women <-
  logit_post_menop_women %>% 
  tidy %>% 
  mutate(OR = exp(estimate),
         group = "Post_menopause_women")
```

### Diagnostics

We check that there is no problematic collinearity betwen regressors (VIF < 5)

```{r}
vif(logit_post_menop_women)
```

```{r}
plot(logit_post_menop_women)
```




## Men

```{r}
regression_data_men <-
  regression_data %>% 
  filter(group == "Men") %>% 
  dplyr::select(donor, age, starting_weight, BMI_diff,
                log_Ferritin_beginning, Hb_beginning, 
                donation_frequency) %>% 
  gather(key = key, value = value, -donor) %>% 
  group_by(key) %>% 
  mutate(value_scale = scale(value, scale = FALSE)[,1]) %>% 
  dplyr::select(-value) %>% 
  spread(key = key, value = value_scale) %>% 
  inner_join(regression_data %>% 
               dplyr::select(donor, health_poor,health_satisfactory, health_very_good, health_excellent,health_outcome,health_outcome_neg,smoking_behavior),
             by = "donor")
```


### Correlograms

```{r echo=FALSE}

ggpairs(regression_data_men, 
        columns = c("health_outcome", "age", "starting_weight",  "BMI_diff", "log_Ferritin_beginning", "Hb_beginning", "donation_frequency"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)
```

### Models and results

```{r}
logit_men <- glm(health_outcome ~ age + starting_weight + BMI_diff + 
                   log_Ferritin_beginning + Hb_beginning + donation_frequency +
             health_very_good + health_excellent,
             data=regression_data_men,
             family="binomial")
 
 summary(logit_men)
```

```{r}
library(epiDisplay)
logistic.display(logit_men)
```

```{r}
tidy_logit_men <-
  logit_men %>% 
  tidy %>% 
  mutate(OR = exp(estimate),
         group = "Men")
```

### Diagnostics
```{r}
plot(logit_men)
```



```{r}
vif(logit_men)
```


```{r}
reg_data<- list(men=regression_data_men,womenpost=regression_data_post_menop_women,womenpre=regression_data_pre_menop_women)
file <- "../results/hypregdata.rdata"
save(reg_data,file=file)
```


# Bootstraps
```{r}


get_coefficients <- function(data, boot_ind){
  fit <- glm (health_outcome~.,
              data[boot_ind,],
              family = "binomial")
  return(coef(fit))
}



compute_Bca_CI <-function(fit_boot){
  Bca_inf = rep(0, length(fit_boot$t0))
  Bca_sup = rep(0, length(fit_boot$t0))
  for (i_regressor in 1:length(fit_boot$t0)){
    CI <- boot.ci(fit_boot, type = "bca", index=i_regressor)
    Bca_inf[i_regressor] <- CI$bca[4]
    Bca_sup[i_regressor] <- CI$bca[5]
  }
  return(tibble(Bca_inf,Bca_sup, regressor = names(fit_boot$t0)) %>% 
           filter(regressor != "(Intercept)"))
}


get_bootstrap <- function(data, nb_boot){
 
 

    fit_boot <- boot(data, statistic= get_coefficients, R = nb_boot)


  fit_boot_distrib <- as.tibble(fit_boot$t)  
  names(fit_boot_distrib) <- names(fit_boot$t0)

  fit_boot_Bca <- compute_Bca_CI(fit_boot)
  
  return(list(fit_boot_distrib,fit_boot_Bca))
}




get_bootstrap_coeffs <- function(regression_data, current_group, nb_boot )
{

  ## preprocess and standardize data

  
    test_data_std <- regression_data %>% 
      filter(group == current_group) %>% 
      dplyr::select(donor, age, starting_weight, BMI_diff,
                    log_Ferritin_beginning, Hb_beginning, 
                    donation_frequency) %>% 
      gather(key = key, value = value, -donor) %>% 
      group_by(key) %>% 
      mutate(value_scale = scale(value, scale = FALSE)[,1]) %>% 
      dplyr::select(-value) %>% 
      spread(key = key, value = value_scale) %>% 
      inner_join(regression_data %>% 
                   dplyr::select(donor, health_outcome, health_very_good, health_excellent),
                 by = "donor") %>% 
    dplyr::select(-donor) 

  
  
  result <- test_data_std %>% 
  get_bootstrap(nb_boot = nb_boot)


  bootstrap_distrib <- result[[1]] %>% 
    gather(key = regressor, value = coefficient) %>% 
    filter(regressor != "(Intercept)") %>% 
    mutate(group = current_group)
    
  Bca_CI  <- result[[2]] %>% mutate(group = current_group)

  return(list(bootstrap_distrib,Bca_CI))
}


output_file_distrib = ("../results/bootstraps/coeff_boot_distrib_seed_125_ctr_strata_final.rdata")
output_file_Bca = ("../results/bootstraps/coeff_boot_Bca_seed_125_ctr_strata_final.rdata")


#get_from_file_ferr = TRUE

get_from_file_ferr <- all(file.exists(output_file_distrib),file.exists(output_file_Bca) )

set.seed(125)
group_str = "Pre_menopause_women"

if (get_from_file_ferr){
  load(file=output_file_distrib)
  load(file=output_file_Bca)
}else
{
   for(group_str in c("Pre_menopause_women", "Post_menopause_women","Men")){
    
    stuff <- get_bootstrap_coeffs(regression_data, group_str, nb_boot = 10000)
    
    if(!exists("bootstrap_distrib")){
      bootstrap_distrib <- stuff[[1]]
      bootstrap_Bca_CI <- stuff[[2]]
    }else
    {
      bootstrap_distrib<-bind_rows(bootstrap_distrib, stuff[[1]])
      bootstrap_Bca_CI<-bind_rows(bootstrap_Bca_CI, stuff[[2]])
    }
  } 
  save(bootstrap_distrib,file=output_file_distrib) 
  save(bootstrap_Bca_CI,file=output_file_Bca)
}




```

# Forest plots

```{r}
# "\U0394 BMI"

regressor_values <- c(
 "age" = "Age",
   "BMI_diff" = "BMI difference",
   "donation_frequency" = "Donation frequency",
   "Hb_beginning" = "Initial Hemoglobin",
   "log_Ferritin_beginning" ="Initial Ferritin",
   "starting_weight" ="Initial weight",
   "health_very_good" ="Initial health rating \n Very good vs Satisfactory/Good",
    "health_excellent" = "Initial health rating \n Excellent vs Satisfactory/Good"
)
```

```{r}
bootstrap_Bca_CI <-
  tidy_logit_men %>% 
  bind_rows(tidy_logit_post_menop_women) %>% 
  bind_rows(tidy_logit_pre_menop_women) %>% 
  dplyr::select(group, term, OR, p.value) %>% 
  rename(regressor = term) %>% 
  full_join(bootstrap_Bca_CI,
            by =  c("regressor", "group")) %>% 
  filter(regressor != "(Intercept)") %>% 
  mutate(Bca_inf = exp(Bca_inf),
         Bca_sup = exp(Bca_sup),
         is_sig = p.value < 0.05) %>% 
  mutate(regressor = plyr::revalue(regressor, regressor_values),
    regressor = ordered(regressor,
                    levels =  c("Age", "Initial weight", "BMI difference",  "Initial Hemoglobin", 
                                "Initial Ferritin","Donation frequency",
                               "Initial health rating \n Very good vs Satisfactory/Good",
                               "Initial health rating \n Excellent vs Satisfactory/Good")),
    regressor = fct_rev(regressor),
    group = case_when(group == "Pre_menopause_women" ~ "Pre-menopausal women",
      group == "Post_menopause_women" ~ "Post-menopausal women",
      TRUE ~ "Men"),
    group = ordered(group, levels = c(  "Pre-menopausal women",  "Post-menopausal women", "Men"  ) ))

file <- "../results/hyp_bootstrap_Bca_CI.csv"
temp <- bootstrap_Bca_CI %>% mutate_if(is.numeric, round, digits = 3) %>% dplyr::select(regressor,group,p.value,OR,Bca_inf,Bca_sup) %>% arrange(desc(regressor),group)
write.csv(temp ,file = file,row.names = FALSE)

```

```{r , fig.width=18}

# "\U0394 BMI"
bootstrap_distrib %>% 
  mutate(regressor = plyr::revalue(regressor, regressor_values),
  regressor = ordered(regressor,
                  levels =  c("Age", "Initial weight", "BMI difference",  "Initial Hemoglobin", 
                              "Initial Ferritin","Donation frequency",
                              "Initial health rating \n Very good vs Satisfactory/Good",
                              "Initial health rating \n Excellent vs Satisfactory/Good")),
  regressor = fct_rev(regressor),
  group = case_when(group == "Pre_menopause_women" ~ "Pre-menopausal women",
    group == "Post_menopause_women" ~ "Post-menopausal women",
    TRUE ~ "Men"),
  group = ordered(group, levels = c( "Men",  "Post-menopausal women", "Pre-menopausal women" ) )) %>% 
  mutate(OR = exp(coefficient)) %>% 
  ggplot(aes(x = regressor, y = OR, color = group, fill = group)) +
  geom_hline(yintercept = 1 , color = "grey", linetype = "dotted",
             size = 1) +
  # geom_violin(alpha = 0.15, trim = TRUE, position = position_dodge(width = 0.8), size = 0.1) +
  geom_pointrange(data = bootstrap_Bca_CI, 
                  aes(ymin=Bca_inf, ymax = Bca_sup, shape = is_sig, color = group, fill = group),
                  position = position_dodge(width = 0.7), size = 1) +
  # geom_point(data = bootstrap_Bca_CI, 
  #            aes(shape = is_sig, fill = group),
  #            position = position_dodge(width = 0.8), size = 2) +
    scale_color_manual(values = c(  "#0065BD", "#BA0303", "#C50084" ),
                     limits = c("Men",  "Post-menopausal women","Pre-menopausal women" )) +
     scale_fill_manual(values = c(  "#0065BD", "#BA0303", "#C50084" ),
                     limits = c("Men",  "Post-menopausal women","Pre-menopausal women" )) +
 scale_shape_manual(values = c(1,16)) +
  scale_y_log10(limits=c(1/40,40)) +
  # scale_x_discrete(labels=parse(text=unique(bootstrap_Bca_CI$regressor)))
  # ylim(1/35, 35) +
  guides(shape = FALSE) +
  theme_classic() +
  # theme(legend.position = c(0.8,0.9),
  #       legend.title = element_blank(),
  #       legend.box = "vertical",
  #       legend.text = element_text(size = 20),
  #       plot.title =  element_text(size = 24,
  #                                 hjust = 0),
  #       panel.grid.minor.y = element_blank(),
  #       axis.line = element_line(colour="black"),
  #       axis.title.y = element_blank(),
  #       axis.title.x = element_text(size = 24),
  #       axis.text = element_text(size = 22),
  #       panel.grid.major.y = element_blank()) +
  coord_flip() +
  xlab("Robust standardized coefficient") +
  labs(title = NULL,
       subtitle = NULL) +


   theme(axis.title = element_text(size = 14),
        strip.text =  element_text(size = 14),
        legend.text = element_text(size = 12),
        strip.background = element_rect(fill = "white"),
        strip.background.x = element_blank(),
        axis.text.y = element_text(size = 14),
        axis.text.x = element_text(size = 14, angle = 30,hjust = 0.8),
        plot.title =  element_text(size = 16),
        plot.subtitle = element_text(size = 16,
                                     hjust = 0.5))
   #
ggplot2::ggsave(filename = "../results/figures/hyp_log_reg_results_b.pdf",
       width = 35,
       height = 18,
       dpi = 600,
       units = "cm")

```




# Separate plots for individual variables

```{r}
fig.width=15
fig.height=20
```


## Initial hemoglobin (g/l)
```{r}
phbd <- final_data %>% 
  dplyr::select(donor, group, health, questionnaire) %>% 
  drop_na(group) %>% 
  filter( group != "Women_pre_menop_no_mens") %>% 
  spread(key = questionnaire, value = health) %>% 
  filter(First != "NA" & Second != "NA") %>%  
  mutate(Difference = case_when(First < Second ~ "Improved",
                                First == Second ~ "Stable",
                                TRUE ~ "Worsened")) %>% 
    mutate(Difference = ordered(Difference,levels=c("Worsened","Stable","Improved"))) %>% 
  inner_join(FD_activity_biomarker_data, by = c("donor", "group")) %>% 
  mutate(group = case_when(group == "Pre_menopause_women" ~ "Pre-menopausal \n women",
                           group == "Post_menopause_women" ~ "Post-menopausal \n women",
                           TRUE ~ "Men"),
         group = ordered(group, levels = c(  "Men", "Post-menopausal \n women", "Pre-menopausal \n women") ))

phb <-  ggplot(phbd,aes(x = Difference, y = Hb_v, color = group)) +
  # geom_hline(yintercept = 15, color = "red", linetype = 3) +
  geom_boxplot(outlier.alpha = 0) +
  geom_beeswarm(alpha = 0.5, shape = 1) +
  # scale_y_log10() +
  facet_grid(group ~.,scales="free_y") +
  scale_color_manual(values = c( "#ff85a2",  "#de2d26","#00BFFF" ),
                     limits = c("Pre-menopausal \n women", "Post-menopausal \n women","Men" )) +
  theme_bw() +
  theme(legend.position = "none") +
  xlab("Evolution of self-rated health") +
  ylab("Initial hemoglobin (g/l)")  


ggplot2::ggsave(plot=phb,filename = "../results/figures/hyp_hb.pdf",
       width = fig.width,
       height = fig.height,
       dpi = 600,
       units = "cm")

phb
```
```{r}
phbd %>% group_by(group, Difference) %>% summarise(mean=mean(Hb_v),median=median(Hb_v))

```

```{r}
phbd %>% group_by(group) %>% summarise(mean=mean(Hb_v),median=median(Hb_v))

```



## Initial ferritin
```{r}
pferd <- final_data %>% 
  dplyr::select(donor, group, health, questionnaire) %>% 
  drop_na(group) %>% 
  filter( group != "Women_pre_menop_no_mens") %>% 
  spread(key = questionnaire, value = health) %>% 
  filter(First != "NA" & Second != "NA") %>%  
  mutate(Difference = case_when(First < Second ~ "Improved",
                                First == Second ~ "Stable",
                                TRUE ~ "Worsened")) %>% 
    mutate(Difference = ordered(Difference,levels=c("Worsened","Stable","Improved"))) %>% 
  inner_join(FD_activity_biomarker_data, by = c("donor", "group")) %>% 
  mutate(group = case_when(group == "Pre_menopause_women" ~ "Pre-menopausal \n women",
                           group == "Post_menopause_women" ~ "Post-menopausal \n women",
                           TRUE ~ "Men"),
         group = ordered(group, levels = c(  "Men", "Post-menopausal \n women", "Pre-menopausal \n women") ))
  
pfer <-  ggplot(pferd, aes(x = Difference, y = Ferritin_beginning, color = group)) +
  # geom_hline(yintercept = 15, color = "red", linetype = 3) +
  geom_boxplot(outlier.alpha = 0) +
  geom_beeswarm(alpha = 0.5, shape = 1) +
   scale_y_log10() +
  facet_grid(group ~.,scales="free_y") +
  #facet_grid(group ~.) +
  scale_color_manual(values = c( "#ff85a2",  "#de2d26","#00BFFF" ),
                     limits = c("Pre-menopausal \n women", "Post-menopausal \n women","Men" )) +
  theme_bw() +
  theme(legend.position = "none") +
  xlab("Evolution of self-rated health") +
  ylab("Initial ferritin (ug/l)") 
  #scale_y_log10(limits=c(10,100)) 
  
  


ggplot2::ggsave(plot=pfer,filename = "../results/figures/hyp_ferritin.pdf",
       width = fig.width,
       height = fig.height,
       dpi = 600,
       units = "cm")

pfer
```
```{r}
pferd %>% group_by(group, Difference) %>% summarise(mean=mean(Ferritin_beginning),median=median(Ferritin_beginning),geoMean=exp(mean(log(Ferritin_beginning))))

```

```{r}
pferd %>% group_by(group) %>% summarise(mean=mean(Ferritin_beginning),median=median(Ferritin_beginning),geoMean=exp(mean(log(Ferritin_beginning))))

```



## BMI difference
```{r}
pbmd <- final_data %>% 
  dplyr::select(donor, group, health, questionnaire) %>% 
  drop_na(group) %>% 
  filter( group != "Women_pre_menop_no_mens") %>% 
  spread(key = questionnaire, value = health) %>% 
  filter(First != "NA" & Second != "NA") %>%  
  mutate(Difference = case_when(First < Second ~ "Improved",
                                First == Second ~ "Stable",
                                TRUE ~ "Worsened")) %>%
  mutate(Difference = ordered(Difference,levels=c("Worsened","Stable","Improved"))) %>% 
  inner_join(FD_activity_biomarker_data, by = c("donor", "group")) %>% 
  mutate(group = case_when(group == "Pre_menopause_women" ~ "Pre-menopausal \n women",
                           group == "Post_menopause_women" ~ "Post-menopausal \n women",
                           TRUE ~ "Men"),
         group = ordered(group, levels = c(  "Men", "Post-menopausal \n women", "Pre-menopausal \n women") ))

pbm <-  ggplot(pbmd, aes(x = Difference, y = BMI_diff, color = group)) +
  # geom_hline(yintercept = 15, color = "red", linetype = 3) +
  geom_boxplot(outlier.alpha = 0) +
  geom_beeswarm(alpha = 0.5, shape = 1) +
  #scale_y_log10() +
  facet_grid(group ~.,scales="free_y") +
  scale_color_manual(values = c( "#ff85a2",  "#de2d26","#00BFFF" ),
                     limits = c("Pre-menopausal \n women", "Post-menopausal \n women","Men" )) +
  theme_bw() +
  theme(legend.position = "none") +
  xlab("Evolution of self-rated health") +
  ylab("BMI difference") 
  #ylim(-5,5)

ggplot2::ggsave(plot=pbm,filename = "../results/figures/hyp_bmi_diff.pdf",
       width = fig.width,
       height = fig.height,
       dpi = 600,
       units = "cm")

pbm
```
```{r}
pbmd %>% group_by(group, Difference) %>% summarise(mean=mean(BMI_diff),median=median(BMI_diff))

```

```{r}
pbmd %>% group_by(group) %>% summarise(mean=mean(BMI_diff),median=median(BMI_diff))

```


## Donation frequency
```{r}
pfqd <- final_data %>% 
  dplyr::select(donor, group, health, questionnaire) %>% 
  drop_na(group) %>% 
  filter( group != "Women_pre_menop_no_mens") %>% 
  spread(key = questionnaire, value = health) %>% 
  filter(First != "NA" & Second != "NA") %>%  
  mutate(Difference = case_when(First < Second ~ "Improved",
                                First == Second ~ "Stable",
                                TRUE ~ "Worsened")) %>% 
    mutate(Difference = ordered(Difference,levels=c("Worsened","Stable","Improved"))) %>% 
  inner_join(FD_activity_biomarker_data, by = c("donor", "group")) %>%
  mutate(donation_frequency = nb_donations_between_questionnaires/(interval_between_quest/12) ) %>% 
  mutate(group = case_when(group == "Pre_menopause_women" ~ "Pre-menopausal \n women",
                           group == "Post_menopause_women" ~ "Post-menopausal \n women",
                           TRUE ~ "Men"),
         group = ordered(group, levels = c(  "Men", "Post-menopausal \n women", "Pre-menopausal \n women") )) 
pfq <-  ggplot(pfqd, aes(x = Difference, y = donation_frequency, color = group)) +
  # geom_hline(yintercept = 15, color = "red", linetype = 3) +
  geom_boxplot(outlier.alpha = 0) +
  geom_beeswarm(alpha = 0.5, shape = 1) +
  #scale_y_log10() +
  facet_grid(group ~.,scales="free_y") +
  scale_color_manual(values = c( "#ff85a2",  "#de2d26","#00BFFF" ),
                     limits = c("Pre-menopausal \n women", "Post-menopausal \n women","Men" )) +
  theme_bw() +
  theme(legend.position = "none") +
  xlab("Evolution of self-rated health") +
  ylab("Doonation frequency (yearly)")
  


ggplot2::ggsave(plot=pfq,filename = "../results/figures/hyp_n_donations.pdf",
       width = fig.width,
       height = fig.height,
       dpi = 600,
       units = "cm")

pfq
```
```{r}
pfqd %>% group_by(group, Difference) %>% summarise(mean=mean(donation_frequency),median=median(donation_frequency))

```

```{r}
pfqd %>% group_by(group) %>% summarise(mean=mean(donation_frequency),median=median(donation_frequency))

```


## Inital weigth
```{r}
pwed <- final_data %>% 
  dplyr::select(donor, group, health, questionnaire) %>% 
  drop_na(group) %>% 
  filter( group != "Women_pre_menop_no_mens") %>% 
  spread(key = questionnaire, value = health) %>% 
  filter(First != "NA" & Second != "NA") %>%  
  mutate(Difference = case_when(First < Second ~ "Improved",
                                First == Second ~ "Stable",
                                TRUE ~ "Worsened")) %>% 
    mutate(Difference = ordered(Difference,levels=c("Worsened","Stable","Improved"))) %>% 
  inner_join(FD_activity_biomarker_data, by = c("donor", "group")) %>% 
  mutate(group = case_when(group == "Pre_menopause_women" ~ "Pre-menopausal \n women",
                           group == "Post_menopause_women" ~ "Post-menopausal \n women",
                           TRUE ~ "Men"),
         group = ordered(group, levels = c(  "Men", "Post-menopausal \n women", "Pre-menopausal \n women") )) 
pwe <-  ggplot(pwed, aes(x = Difference, y = starting_weight, color = group)) +
  # geom_hline(yintercept = 15, color = "red", linetype = 3) +
  geom_boxplot(outlier.alpha = 0) +
  geom_beeswarm(alpha = 0.5, shape = 1) +
  scale_y_log10() +
  facet_grid(group ~.,scales="free_y") +
  scale_color_manual(values = c( "#ff85a2",  "#de2d26","#00BFFF" ),
                     limits = c("Pre-menopausal \n women", "Post-menopausal \n women","Men" )) +
  theme_bw() +
  theme(legend.position = "none") +
  xlab("Evolution of self-rated health") +
  ylab("Initial weight (kg)") 
  
  


ggplot2::ggsave(plot=pwe,filename = "../results/figures/hyp_weight.pdf",
       width = fig.width,
       height = fig.height,
       dpi = 600,
       units = "cm")

pwe
```


```{r}
pwed %>% group_by(group, Difference) %>% summarise(mean=mean(starting_weight),median=median(starting_weight))

```

```{r}
pwed %>% group_by(group) %>% summarise(mean=mean(starting_weight),median=median(starting_weight))

```
