---
title: "Inital data analysis for FINDONOR- CHANGES IN SELF-RATED HEALTH DURING THE STUDY AND IT'S RELATION TO IRON BIOMARKERS "
author: "Mikko Arvas"
date: "`r Sys.time()`"
output: html_document
---


```{r setup, include=FALSE}
Sys.setlocale("LC_MESSAGES", 'en_GB.UTF-8')
Sys.setenv(LANG = "en_US.UTF-8")
# echo "rmarkdown::render('inital_data_analysis.Rmd', clean=TRUE,output_format='html_document',output_file='../results/inital_data_analysis.html')" | R --slave
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
library(tidyverse)
library(corrplot)
library(knitr)
library(lubridate)
```



# Executive summary


In this file we carry out inital data analysis to decide which explenatory variables to exclude based on correlations between the explenatory variables, shapes of variables distributions and counts of missing data. Then the chosen variables are coded as change between the 2 questionnaires.

# Load the data

We load the data and assign groups as they were assigned in the first part of the study. 

```{r loading}
#Load questionnaire data
load(file = "../data/final_data.rdata") 

data %>% group_by(questionnaire) %>% count()

```

```{r}
data %>% group_by(questionnaire,group) %>% filter(group != 'Women_pre_menop_no_mens') %>%  count()
```


#Remove end point variable

In the inital data analysis stage we do not want to look at our primary endpoint: "self-rated health" i.e. "health" to avoid fishing for correlations.
```{r}

data <- data %>% filter(!is.na(health))  %>% select(-health)
data %>% group_by(questionnaire) %>% count()
```


#Preprocessing recodings
Questions with 3 possible answers : "no", "yes", "don't know", are recoded as an ordered factor with 3 levels: "yes"< "don't know" <  "no".  
```{r}
#Have you felt that you were able to accomplish less than you would have liked to?
temp <- as.character(data$accomplish)
temp[is.na(temp)] <- "don't know"
data$accomplish = ordered(temp,
                levels =  c("yes","don't know", "no"))
#During the past four weeks, has your mental health interfered with your work or other daily activities?
temp <- as.character(data$ment_inf)
temp[is.na(temp)] <- "don't know"
data$ment_inf = ordered(temp,
                levels =  c("yes","don't know", "no"))

#During the past four weeks, have your physical symptoms (such as pain) or health in general interfered with your work or other daily activities?
temp <- as.character(data$phys_inf)
temp[is.na(temp)] <- "don't know"
data$phys_inf = ordered(temp,
                levels =  c("yes","don't know", "no"))



```



# How much data do we have
```{r}
#drop out variables that, 
# we know from our previous work are too messy: iron_supp_detailed
# were only asked in second questionaire: life_qual, education, employment
# are conveniency variables: AgeGroup, SixMonthsFromStartCount_FB, TwoYearsFromStartCount_FB
# is ID: donor
temp <- data %>% select(-iron_supp_detailed, -life_qual, -education, -employment,-AgeGroup, -SixMonthsFromStartCount_FB,-TwoYearsFromStartCount_FB,-donor)
temp <- na.omit(temp)
temp %>% group_by(questionnaire) %>% count()
```

From `r ncol(temp)` questions.

As we will loose one third of our data if we study only complete data we have try to make decisions on which questions to exclude prior to analyses or which data to impute.

```{r}
temp <- data %>% select(-iron_supp_detailed, -life_qual, -education, -employment,-AgeGroup, -SixMonthsFromStartCount_FB,-TwoYearsFromStartCount_FB,-donor)
temp <- temp %>% group_by(questionnaire) %>%  summarise_all(funs(sum(is.na(.))))
temp <- as_tibble(t(temp), rownames = "Variable") %>% filter(Variable != "questionnaire")
colnames(temp)[2] <- "First"
colnames(temp)[3] <- "Second"
temp <- temp %>%  gather(key="questionnaire",value="NAs",-Variable) %>% mutate(NAs = as.numeric(gsub(" ","",NAs))) %>% filter(NAs != 0)

ggplot(temp) + geom_col(aes(x=Variable,y=NAs,fill=questionnaire),pos="dodge") +  coord_flip() 

```



```{r}
temp <- temp %>% arrange(NAs) %>% tail(10)

ggplot(temp) + geom_col(aes(x=Variable,y=NAs,fill=questionnaire),pos="dodge") +  coord_flip() 

```

People cannot remember if they have taken vitamine supplements containing iron or vitamine C. Let's first see if some of these will drop out because they correlate heavily with some other explenatory variable.

# Correlations between variables

```{r , warning=FALSE}

temp <- data %>% select(-iron_supp_detailed, -life_qual, -education, -employment,-AgeGroup, -SixMonthsFromStartCount_FB,-TwoYearsFromStartCount_FB, -donor,-questionnaire,-sex,-group)
temp$date <- as.numeric(temp$date)

#Change the numeric to factors to enable similar correlation calculation for all
temp <- temp %>% mutate_if(is.numeric, ~cut(.,5))

#https://datascience.stackexchange.com/questions/893/how-to-get-correlation-between-two-categorical-variable-and-a-categorical-variab
cormat <- matrix(NA,ncol = ncol(temp),nrow=ncol(temp))
rownames(cormat) <-colnames(temp)
colnames(cormat) <-colnames(temp)
for (i in rownames(cormat)) {
  for (j in colnames(cormat)) {
    d <- table(as.factor(temp[[i]]),as.factor(temp[[j]]))
#    print(d)
    #drop any all zero rows
    d <- d[apply(X=d,MARGIN = 1,FUN = function(x){!all(x == 0 )}),]
    d <- d[,apply(X=d,MARGIN = 2,FUN = function(x){!all(x == 0 )})]
    #print(d)
    t <- chisq.test(d,correct = F)
    cramersV <- sqrt(t$statistic / sum(d))
    if (cramersV > 1) {
      #sometimes it fails in perfect correlation
      cramersV <- 1
    }
    cormat[i,j] <- cramersV
  }
}


```

Visualise correlation disribution
```{r}
temp <- data.frame(correlation=cormat[lower.tri(cormat)])
ggplot(temp) +  geom_histogram(aes(x=correlation),binwidth = 0.1)

#and create a fake significance matrix of arbitary correlation of 0.5
p.mat <- matrix(1, nrow=nrow(cormat),ncol=ncol(cormat))
p.mat[as.vector(cormat > 0.5)] <- 0.01 
diag(p.mat) <- 1

```


```{r}
corrplot(cormat,type="lower",order="hclust",p.mat=p.mat,sig.level = 0.05,insig = "label_sig",pch.col = 'red',pch.cex = 1)
```

Extract those with correlation above 0.5.
```{r}
#take anything with at least 0.5 correlation
any_above <- function(x){any(x > 0.5) }
temp <-cormat
#empty the diagonal
diag(temp) <- 0
#filter fake significance matrix
p.mat <- p.mat[apply(temp,1, any_above),apply(temp,2, any_above)]
#filter correlation matrix
temp <- temp[apply(temp,1, any_above),apply(temp,2, any_above)]
p.mat <- p.mat[apply(temp,1, any_above),apply(temp,2, any_above)]
corrplot(temp,type="lower",order="hclust",p.mat=p.mat,sig.level = 0.05,insig = "label_sig",pch.col = 'red',pch.cex = 1)
```

#Choose what to keep
Between those that correlate above 0.5 choose which variable to keep to present the correlaiting variables in the exploratory analysis.

## act_light, act_heavy
```{r}
ggplot(data) + geom_bar(aes(x=act_light))
```

```{r}
ggplot(data) + geom_bar(aes(x=act_heavy))
```

```{r}
table(is.na(data$act_heavy))
table(is.na(data$act_light))
```
act_heavy looses more data, but has much more spread pontentially contaning more information.

Keep "act_heavy". 

##health_c vs phys_cond

```{r}
ggplot(data) + geom_bar(aes(x=health_c))
```
```{r}
ggplot(data) + geom_bar(aes(x=phys_cond))
```


```{r}
table(is.na(data$health_c))
table(is.na(data$phys_cond))
```


phys_cond looks better and looses less data. 

##vita_c vs vita_multi

```{r}
ggplot(data) + geom_bar(aes(x=vita_C))
```

```{r}
ggplot(data) + geom_bar(aes(x=vita_multi))
```

```{r}
table(is.na(data$vita_C))
table(is.na(data$vita_multi))
```


vita_multi looses much less data and looks better so vita_C can be dropped.

##sleep_night vs sleep_24h
```{r , warning=FALSE}
ggplot(data) + geom_histogram(aes(x=sleep_night),binwidth = 1)
```
```{r , warning=FALSE}
ggplot(data) + geom_histogram(aes(x=sleep_24h),binwidth = 1)
```

```{r}
table(is.na(data$sleep_night))
table(is.na(data$sleep_24h))
```

Distributions look quite similar, but sleep_night looses less data. 

##freq_fruits vs freq_vege
```{r , warning=FALSE}
ggplot(data) + geom_bar(aes(x=freq_fruits))
```
```{r , warning=FALSE}
ggplot(data) + geom_bar(aes(x=freq_vege))
```

```{r}
table(is.na(data$freq_fruits))
table(is.na(data$freq_vege))
```

freq_fruits look better and looses less data.

##freq_sports vs phys_cond
```{r , warning=FALSE}
ggplot(data) + geom_bar(aes(x=freq_sports))
```
```{r , warning=FALSE}
ggplot(data) + geom_bar(aes(x=phys_cond))
```

```{r}
table(is.na(data$freq_sports))
table(is.na(data$phys_cond))
```

Hard to say anything about the distributions, but physical condition looses less data.

##freq_spirits, freq_beer, freq_wine
```{r , warning=FALSE}
ggplot(data) + geom_bar(aes(x=freq_spirits))
```
```{r , warning=FALSE}
ggplot(data) + geom_bar(aes(x=freq_beer))
```
```{r , warning=FALSE}
ggplot(data) + geom_bar(aes(x=freq_wine))
```
```{r}
table(is.na(data$freq_spirits))
table(is.na(data$freq_beer))
table(is.na(data$freq_wine))
```
Hard to say, but freq_beer looses least data. But this does not cover at all persons that drink mostly cider which could be in particular found in pre-menopausal women.

To have have some kind of proxy over these incomplete but overlapping questions we will create a new variable freq_alc.

```{r}
#Combine the alcohol questions
data$freq_alc <- as.ordered(apply(cbind(data$freq_beer,data$freq_wine,data$freq_spirits),1,max,na.rm=TRUE))
#Make it in to a factor
levels(data$freq_alc) <- levels(data$freq_beer)

```


##freq_cutlets, freq_meat, diet, freq_meat, freq_fish, freq_eggs
```{r , warning=FALSE}
ggplot(data) + geom_bar(aes(x=freq_cutlets))
```
```{r , warning=FALSE}
ggplot(data) + geom_bar(aes(x=freq_red_meat))
```
```{r , warning=FALSE}
ggplot(data) + geom_bar(aes(x=diet))
```
```{r , warning=FALSE}
ggplot(data) + geom_bar(aes(x=freq_fish))
```
```{r , warning=FALSE}
ggplot(data) + geom_bar(aes(x=freq_eggs))
```

```{r}
table(is.na(data$freq_cutlets))
table(is.na(data$freq_red_meat))
table(is.na(data$diet))
table(is.na(data$freq_fish))
table(is.na(data$freq_eggs))
```

As already seen in the enrollment data paper freq_red_meats distribution looks best also it looses leas data.


##BMI, height and weigth
```{r , warning=FALSE}
ggplot(data) + geom_histogram(aes(x=BMI),binwidth = 1)
```
```{r , warning=FALSE}
ggplot(data) + geom_histogram(aes(x=height),binwidth = 5)
```

```{r , warning=FALSE}
ggplot(data) + geom_histogram(aes(x=weight),binwidth = 5)
```

```{r}
table(is.na(data$BMI))
table(is.na(data$height))
table(is.na(data$weight))
```

Keep BMI as composite of both.

##freq_energy, freq_sad, health_soc, freq_calm, accomplish
```{r , warning=FALSE}
ggplot(data) + geom_bar(aes(x=freq_energy))
```
```{r , warning=FALSE}
ggplot(data) + geom_bar(aes(x=freq_sad))
```
```{r , warning=FALSE}
ggplot(data) + geom_bar(aes(x=health_soc))
```
```{r , warning=FALSE}
ggplot(data) + geom_bar(aes(x=freq_calm))
```
```{r , warning=FALSE}
ggplot(data) + geom_bar(aes(x=accomplish))
```

```{r}
table(is.na(data$freq_energy))
table(is.na(data$freq_sad))
table(is.na(data$health_soc))
table(is.na(data$freq_calm))
table(is.na(data$accomplish))
```

accomplish looses a lot of data it can dropped just for that.
freq_energy correlates with phys_cond and can be dropped for that.
health_soc has a very skewed distiribution.
freq_calm looses least data and has maybe the best looking distribution.

##h_act_light vs h_act_noen
```{r , warning=FALSE}
ggplot(data) + geom_bar(aes(x=h_act_light))
```
```{r , warning=FALSE}
ggplot(data) + geom_bar(aes(x=h_act_none))
```

```{r}
table(is.na(data$h_act_light))
table(is.na(data$h_act_none))
```

h_act_light looses less data.

#Drop questions based on correlation analysis
Based on the above we have decided to drop: act_light, health_c, vita_C, sleep_24h, freq_vege, freq_sports, freq_spirits, freq_beer, freq_wine, freq_cutlets, diet, freq_fish, freq_eggs, weight, height, accomplish, freq_energy, health_soc, freq_sad, h_act_none

```{r}
temp <- data %>% select(-iron_supp_detailed, -life_qual, -education, -employment,-AgeGroup, -SixMonthsFromStartCount_FB,-TwoYearsFromStartCount_FB, -act_light, -health_c, -vita_C, -sleep_24h, -freq_vege, -freq_sports, -freq_spirits, -freq_wine, -freq_beer, -freq_cutlets, -diet, -freq_fish, -freq_eggs, -weight, -height, -accomplish, -freq_energy, -health_soc, -freq_sad,-h_act_none)
temp <- na.omit(temp)
temp %>% group_by(questionnaire) %>% count()
```
From `r ncol(temp)` questions.

```{r}

#But questionnaire, donor, group, age, sex and date are not procesed into change between questionnaires
tx <- setdiff(colnames(temp),c("questionnaire", "donor", "group", "age", "sex","date" ))
tx
```


```{r}
length(tx)
```

How will the sex/menopause groups look
```{r}
temp %>% group_by(questionnaire,group) %>% filter(group != 'Women_pre_menop_no_mens') %>%  count()
```

By excluding several questions that correlate with other questions and have distributions that are skewed (e.g. diet) or are uniformative (e.g. act_ligth)  or are not clearly ordered factors (like diet) we get complete data for about hundred more persons. There might still be too many explenatory variables for the smallest group i.e. Post_menopause_women and we will still loose data when we look for complete cases in both questionaires.

#Recode answers as change between the questionnaires

Exclude questions that are dealt with later in the regression data processing:
BMI, smoking, health, date, age, group. Health has been already excluded.
```{r}
temp <- temp %>% select(-BMI,-smoking,-date,-age,-group,-sex)
#Extract complete data for both questinaires and separate questionaire answers into different columns for processing
splitbyq <- split(temp,f=temp$questionnaire)
first <- splitbyq[["First"]]
second <- splitbyq[["Second"]]
colnames(second) <- paste0(colnames(second),"_2")
temp <- inner_join(first,second,by=c("donor"="donor_2"))
cat(nrow(temp),"complete cases in both questionnaires\n")
```



```{r}

classes <- sort(unlist(lapply(lapply(first,class),paste,collapse=" ")))
classes <- as.tibble(cbind(names(classes),classes))
no_answers <- first %>% summarise_all(~length(levels(.)))
no_answers <- as_tibble(cbind(variable = names(no_answers), t(no_answers)))
no_answers <- full_join(no_answers,classes,by=c("variable"="V1")) %>% rename(no_answers=V2)
no_answers %>% group_by(no_answers)  %>% count() %>%  arrange(n)
```

```{r}
no_answers %>%  filter(no_answers == 0)
```

no_aswers == 0 are non-factoral variables. Change in cholesterol, blood_sugar, HBP, low_hb_hist, anemia_hist, high_hb_hist is coded as if the status changed during the study.

```{r}
temp <- temp %>% mutate(cholesterol_d = case_when(cholesterol == 0 & cholesterol_2 == 1 ~ 1,
                                          TRUE ~ 0
                                          )) %>% mutate(cholesterol_d = as.logical(cholesterol_d))

temp <- temp %>% mutate(blood_sugar_d = case_when(blood_sugar == 0 & blood_sugar_2 == 1 ~ 1,
                                          TRUE ~ 0
                                          )) %>% mutate(blood_sugar_d = as.logical(blood_sugar_d))

temp <- temp %>% mutate(HBP_d = case_when(HBP == 0 & HBP_2 == 1 ~ 1,
                                          TRUE ~ 0
                                          )) %>% mutate(HBP_d = as.logical(HBP_d))

temp <- temp %>% mutate(low_hb_hist_d = case_when(low_hb_hist == 0 & low_hb_hist_2 == 1 ~ 1,
                                          TRUE ~ 0
                                          )) %>% mutate(low_hb_hist_d = as.logical(low_hb_hist_d))

temp <- temp %>% mutate(anemia_hist_d = case_when(anemia_hist == 0 & anemia_hist_2 == 1 ~ 1,
                                          TRUE ~ 0
                                          )) %>% mutate(anemia_hist_d = as.logical(anemia_hist_d))

temp <- temp %>% mutate(high_hb_hist_d = case_when(high_hb_hist == 0 & high_hb_hist_2 == 1 ~ 1,
                                          TRUE ~ 0
                                          )) %>% mutate(high_hb_hist_d = as.logical(high_hb_hist_d))
```


iron_supp recoded as either reducing, stable or increasing.

```{r}
temp <- temp %>% mutate(iron_supp_d = case_when(iron_supp == FALSE & iron_supp_2 == TRUE ~ "Increased",
                                                iron_supp == TRUE & iron_supp_2 == FALSE ~ "Reduced",
                                          TRUE ~ "Stable"
                                          )) %>% mutate( iron_supp_d= factor(iron_supp_d,levels = c("Reduced","Stable","Increased"),ordered = TRUE))

```

sleep_night recoded as difference

```{r}
temp <- temp %>% mutate(sleep_night_d = sleep_night_2 - sleep_night)
```

```{r}
no_answers %>%  filter(no_answers == 6)
```

These are all ordered factors. They are recoded as  either reducing, stable or increasing.

```{r}
temp <- temp %>% mutate(freq_calm_d = as.numeric(freq_calm_2)  - as.numeric(freq_calm)) %>% 
  mutate(freq_calm_d = case_when(
    freq_calm_d > 0 ~ "Increased",
    freq_calm_d < 0 ~ "Reduced",
    TRUE ~ "Stable"
  )) %>% mutate( freq_calm_d= factor(freq_calm_d,levels = c("Reduced","Stable","Increased"),ordered = TRUE))

temp <- temp %>% mutate(freq_red_meat_d = as.numeric(freq_red_meat_2)  - as.numeric(freq_red_meat)) %>% 
  mutate(freq_red_meat_d = case_when(
    freq_red_meat_d > 0 ~ "Increased",
    freq_red_meat_d < 0 ~ "Reduced",
    TRUE ~ "Stable"
  )) %>% mutate( freq_red_meat_d= factor(freq_red_meat_d,levels = c("Reduced","Stable","Increased"),ordered = TRUE))

temp <- temp %>% mutate(freq_fruits_d = as.numeric(freq_fruits_2)  - as.numeric(freq_fruits)) %>% 
  mutate(freq_fruits_d = case_when(
    freq_fruits_d > 0 ~ "Increased",
    freq_fruits_d < 0 ~ "Reduced",
    TRUE ~ "Stable"
  )) %>% mutate( freq_fruits_d= factor(freq_fruits_d,levels = c("Reduced","Stable","Increased"),ordered = TRUE))

temp <- temp %>% mutate(freq_juices_d = as.numeric(freq_juices_2)  - as.numeric(freq_juices)) %>% 
  mutate(freq_juices_d = case_when(
    freq_juices_d > 0 ~ "Increased",
    freq_juices_d < 0 ~ "Reduced",
    TRUE ~ "Stable"
  )) %>% mutate( freq_juices_d= factor(freq_juices_d,levels = c("Reduced","Stable","Increased"),ordered = TRUE))

temp <- temp %>% mutate(freq_wholem_d = as.numeric(freq_wholem_2)  - as.numeric(freq_wholem)) %>% 
  mutate(freq_wholem_d = case_when(
    freq_wholem_d > 0 ~ "Increased",
    freq_wholem_d < 0 ~ "Reduced",
    TRUE ~ "Stable"
  )) %>% mutate( freq_wholem_d= factor(freq_wholem_d,levels = c("Reduced","Stable","Increased"),ordered = TRUE))

temp <- temp %>% mutate(freq_milk_d = as.numeric(freq_milk_2)  - as.numeric(freq_milk)) %>% 
  mutate(freq_milk_d = case_when(
    freq_milk_d > 0 ~ "Increased",
    freq_milk_d < 0 ~ "Reduced",
    TRUE ~ "Stable"
  )) %>% mutate( freq_milk_d= factor(freq_milk_d,levels = c("Reduced","Stable","Increased"),ordered = TRUE))

temp <- temp %>% mutate(freq_cof_d = as.numeric(freq_cof_2)  - as.numeric(freq_cof)) %>% 
  mutate(freq_cof_d = case_when(
    freq_cof_d > 0 ~ "Increased",
    freq_cof_d < 0 ~ "Reduced",
    TRUE ~ "Stable"
  )) %>% mutate( freq_cof_d= factor(freq_cof_d,levels = c("Reduced","Stable","Increased"),ordered = TRUE))

temp <- temp %>% mutate(freq_tea_d = as.numeric(freq_tea_2)  - as.numeric(freq_tea)) %>% 
  mutate(freq_tea_d = case_when(
    freq_tea_d > 0 ~ "Increased",
    freq_tea_d < 0 ~ "Reduced",
    TRUE ~ "Stable"
  )) %>% mutate( freq_tea_d= factor(freq_tea_d,levels = c("Reduced","Stable","Increased"),ordered = TRUE))

temp <- temp %>% mutate(phys_cond_d = as.numeric(phys_cond_2)  - as.numeric(phys_cond)) %>% 
  mutate(phys_cond_d = case_when(
    phys_cond_d > 0 ~ "Increased",
    phys_cond_d < 0 ~ "Reduced",
    TRUE ~ "Stable"
  )) %>% mutate( phys_cond_d= factor(phys_cond_d,levels = c("Reduced","Stable","Increased"),ordered = TRUE))

```




```{r}
no_answers %>%  filter(no_answers == 4)
```


These are all ordered factors. They are recoded as  either reducing, stable or increasing.
```{r}
temp <- temp %>% mutate(act_heavy_d = as.numeric(act_heavy_2)  - as.numeric(act_heavy)) %>% 
  mutate(act_heavy_d = case_when(
    act_heavy_d > 0 ~ "Increased",
    act_heavy_d < 0 ~ "Reduced",
    TRUE ~ "Stable"
  )) %>% mutate( act_heavy_d= factor(act_heavy_d,levels = c("Reduced","Stable","Increased"),ordered = TRUE))

#sleep_enough
temp <- temp %>% mutate(sleep_enough_d = as.numeric(sleep_enough_2)  - as.numeric(sleep_enough)) %>% 
  mutate(sleep_enough_d = case_when(
    sleep_enough_d > 0 ~ "Increased",
    sleep_enough_d < 0 ~ "Reduced",
    TRUE ~ "Stable"
  )) %>% mutate( sleep_enough_d= factor(sleep_enough_d,levels = c("Reduced","Stable","Increased"),ordered = TRUE))

#day_act
temp <- temp %>% mutate(day_act_d = as.numeric(day_act_2)  - as.numeric(day_act)) %>% 
  mutate(day_act_d = case_when(
    day_act_d > 0 ~ "Increased",
    day_act_d < 0 ~ "Reduced",
    TRUE ~ "Stable"
  )) %>% mutate( day_act_d= factor(day_act_d,levels = c("Reduced","Stable","Increased"),ordered = TRUE))

#h_act_light
temp <- temp %>% mutate(h_act_light_d = as.numeric(h_act_light_2)  - as.numeric(h_act_light)) %>% 
  mutate(h_act_light_d = case_when(
    h_act_light_d > 0 ~ "Increased",
    h_act_light_d < 0 ~ "Reduced",
    TRUE ~ "Stable"
  )) %>% mutate( h_act_light_d= factor(h_act_light_d,levels = c("Reduced","Stable","Increased"),ordered = TRUE))

#vita_multi
temp <- temp %>% mutate(vita_multi_d = as.numeric(vita_multi_2)  - as.numeric(vita_multi)) %>% 
  mutate(vita_multi_d = case_when(
    vita_multi_d > 0 ~ "Increased",
    vita_multi_d < 0 ~ "Reduced",
    TRUE ~ "Stable"
  )) %>% mutate( vita_multi_d= factor(vita_multi_d,levels = c("Reduced","Stable","Increased"),ordered = TRUE))

#vita_iron
temp <- temp %>% mutate(vita_iron_d = as.numeric(vita_iron_2)  - as.numeric(vita_iron)) %>% 
  mutate(vita_iron_d = case_when(
    vita_iron_d > 0 ~ "Increased",
    vita_iron_d < 0 ~ "Reduced",
    TRUE ~ "Stable"
  )) %>% mutate( vita_iron_d= factor(vita_iron_d,levels = c("Reduced","Stable","Increased"),ordered = TRUE))

```

But these 2 are diffent order of time and they are turned around to have same reducing, stable or increasing coding.

sleep_tired Levels: always_or_almost < mostly < rarely < never i.e. is their tiredness reduced, stable or increasing

anti_inf: Levels: daily < not_daily < occasionally < none

```{r}
temp <- temp %>% mutate(sleep_tired_d = as.numeric(sleep_tired)  - as.numeric(sleep_tired_2)) %>% 
  mutate(sleep_tired_d = case_when(
    sleep_tired_d > 0 ~ "Increased",
    sleep_tired_d < 0 ~ "Reduced",
    TRUE ~ "Stable"
  )) %>% mutate( sleep_tired_d= factor(sleep_tired_d,levels = c("Reduced","Stable","Increased"),ordered = TRUE))

temp <- temp %>% mutate(anti_inf_d = as.numeric(anti_inf)  - as.numeric(anti_inf_2)) %>% 
  mutate(anti_inf_d = case_when(
    anti_inf_d > 0 ~ "Increased",
    anti_inf_d < 0 ~ "Reduced",
    TRUE ~ "Stable"
  )) %>% mutate( anti_inf_d= factor(anti_inf_d,levels = c("Reduced","Stable","Increased"),ordered = TRUE))


```


```{r}
no_answers %>%  filter(no_answers == 3)
```

phys_inf Levels: yes < don't know < no -> has your physical health inferred? e.g. first : no, second: yes -> physical health inferring "Increased"

ment_inf Levels: yes < don't know < no -> has your mental health inferred?

```{r}
temp <- temp %>% mutate(phys_inf_d = as.numeric(phys_inf)  - as.numeric(phys_inf_2)) %>% 
  mutate(phys_inf_d = case_when(
    phys_inf_d > 0 ~ "Increased",
    phys_inf_d < 0 ~ "Reduced",
    TRUE ~ "Stable"
  )) %>% mutate( phys_inf_d= factor(phys_inf_d,levels = c("Reduced","Stable","Increased"),ordered = TRUE))

temp <- temp %>% mutate(ment_inf_d = as.numeric(ment_inf)  - as.numeric(ment_inf_2)) %>% 
  mutate(ment_inf_d = case_when(
    ment_inf_d > 0 ~ "Increased",
    ment_inf_d < 0 ~ "Reduced",
    TRUE ~ "Stable"
  )) %>% mutate( ment_inf_d= factor(ment_inf_d,levels = c("Reduced","Stable","Increased"),ordered = TRUE))


```


```{r}
no_answers %>%  filter(no_answers == 5)
```

freq_alc Levels: daily_or_almost < a_few_per_week < a_few_per_month < very_rarely < never 

```{r}
temp <- temp %>% mutate(freq_alc_d = as.numeric(freq_alc)  - as.numeric(freq_alc_2)) %>% 
  mutate(freq_alc_d = case_when(
    freq_alc_d > 0 ~ "Increased",
    freq_alc_d < 0 ~ "Reduced",
    TRUE ~ "Stable"
  )) %>% mutate( freq_alc_d= factor(freq_alc_d,levels = c("Reduced","Stable","Increased"),ordered = TRUE))

```

Select only the change variables and store.

```{r}
cols2keep<- grep("_d$",colnames(temp),value = TRUE)
cols2keep
```
```{r}
cols2keep<- c("donor",cols2keep)
changedata <- temp %>% select(cols2keep)
dim(changedata)
```

#Checks
```{r}
setdiff(colnames(first),gsub("_d$","",colnames(changedata)))
```

```{r}
setdiff(gsub("_d$","",colnames(changedata)),colnames(first))
```


```{r}
classes <- sort(unlist(lapply(lapply(changedata,class),paste,collapse=" ")))
classes <- as.tibble(cbind(names(classes),classes))
levels <- unlist(lapply(lapply(changedata,levels),paste,collapse=" "))
levels <- as.tibble(cbind(names(levels),levels))
changedataclasses <- full_join(classes,levels,by=c("V1"="V1")) %>% rename(variable=V1) 
kable(changedataclasses,caption="Classes and levels of variables only analysed in explorative analysis")
```

```{r}
file <- "../results/changedata.rdata"
save(changedata,file=file)
```

