---
title: "Inital data analysis for FINDONOR- CHANGES IN SELF-RATED HEALTH DURING THE STUDY AND IT'S RELATION TO IRON BIOMARKERS "
author: "Mikko Arvas"
date: "`r Sys.time()`"
output: html_document
---


```{r setup, include=FALSE}
Sys.setlocale("LC_MESSAGES", 'en_GB.UTF-8')
Sys.setenv(LANG = "en_US.UTF-8")
# echo "rmarkdown::render('inital_data_analysis.Rmd', clean=TRUE,output_format='html_document',output_file='../results/inital_data_analysis.html')" | R --slave
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
library(tidyverse)
library(corrplot)
library(knitr)
library(lubridate)
```



# Executive summary


In this file we carry out inital data analysis to decide which explenatory variables to exclude based on correlations between the explenatory variables, shapes of variables distributions and counts of missing data. Then the chosen variables are coded as change between the 2 questionnaires.

# Load the data

We load the data and assign groups as they were assigned in the first part of the study. 

```{r loading}
#Load questionnaire data
load(file = "../data/final_data.rdata") 
data <- data %>% ungroup()
data %>% group_by(questionnaire) %>% count()

```

```{r}
data %>% group_by(questionnaire,group)  %>%  count()
```

```{r}
data <- data %>% filter(group != 'Women_pre_menop_no_mens')
data %>% group_by(questionnaire) %>% count()
```

Here are the questions we are processing. See further details from https://onlinelibrary.wiley.com/action/downloadSupplement?doi=10.1111%2Fvox.12856&file=vox12856-sup-0002-SupInfo2.pdf  

   
Number | Abbreviation | Question text   
-------|--------------|--------------------------------------------------------------------------------------------------   
1.1 |  HBP    |           Have you ever been diagnosed with any of the following conditions? Elevated blood pressure                                                                                              
1.2 |  cholesterol  |     Have you ever been diagnosed with any of the following conditions? Elevated cholesterol levels                                                                                          
1.3 |  blood_sugar    |   Have you ever been diagnosed with any of the following conditions? Elevated blood sugar levels                                                                                            
2   |  low_hb_hist  |     Have you ever been diagnosed with a low haemoglobin (Hb) count?                                                                                                                            
2.1 |  anemia_hist  |     If yes, have you ever been diagnosed with anaemia?                                                                                                                                         
3   |  high_hb_hist |     Have you ever undergone medical tests because of a high haemoglobin count?                                                                                                                
4 |    anti_inf     |     During the past four weeks, have you taken anti-inflammatory pain medication (such as aspirin, ibuprofen or ketoprofen)?                                                                  
5 |    health       |     How would you rate your recent health in general?                                                                                                                                          
6 |    health_c     |     How would you rate your health compared with other people of similar age?                                                                                                                 
7.1 |  act_light    |     Based on your health, how would you rate your ability to carry out the following daily activities? A. Light chores, such as moving a table, vacuuming or cycling.                         
7.2 |  act_heavy    |     Based on your health, how would you rate your ability to carry out the following daily activities? B. Climbing several flights of stairs                                                  
8   |  phys_inf     |     During the past four weeks, have your physical symptoms (such as pain) or health in general interfered with your work or other daily activities?                                          
9   |  ment_inf     |     During the past four weeks, has your mental health interfered with your work or other daily activities?                                                                                   
10  |  accomplish   |     Have you felt that you were able to accomplish less than you would have liked to?                                                                                                         
11.A  | freq_calm   |      During the past four weeks, how often have you felt: calm and relaxed?                                                                                                                   
11.B  | freq_energy |      During the past four weeks, how often have you felt: full of energy?                                                                                                                    
11.C  | freq_sad    |      During the past four weeks, how often have you felt: sad?                                                                                                                                
12  |  health_soc   |     During the past four weeks, how often has your physical or emotional health interfered with your social activities (such as visiting friends or relatives)?                               
13.A  | sleep_night |      How many hours do you sleep on average A. per night (round to the nearest half hour)?                                                                                                    
13.B  | sleep_24h   |      How many hours do you sleep on average B. over a 24-hour period, including nights and daytime naps (round to the nearest half hour)?                                                     
14.A  | sleep_enough |      Sleep habits A. Do you think you get enough sleep?                                                                                                                                      
14.B  | sleep_tired   |    Sleep habits B. Do you feel tired during the day?                                                                                                                                        
15.A  | vita_multi    |    During the past four weeks, have you taken A. multivitamin tablets/supplements?                                                                                                          
15.B  | vita_C        |    During the past four weeks, have you taken B. vitamin C tablets/supplements?                                                                                                           
15.C  | vita_iron     |    During the past four weeks, have you taken C. iron tablets/supplements?                                                                                                                
16    |iron_supp      |   Were you given iron tablets when you last donated blood?                                                                                                                              
16.1  |iron_supp_detailed | If yes, how many iron tablets did you take?                                                                                                                                           
17    | diet          |    Do you follow a special diet?                                                                                                                                                        
18.A  |freq_red_meat  |   How often do you have A. red meat (beef, pork, lamb, game) as the main course?                                                                                                           
18.B  |freq_cutlets|      How often do you have B. cutlets?                                                                                                                                                     
18.C  |freq_fish    |     How often do you have C. fish?                                                                                                                                                        
18.D  |freq_eggs|         How often do you have D. eggs?                                                                                                                                                        
18.E  |freq_fruits |      How often do you have E. fruit and berries?                                                                                                                                           
18.F  |freq_vege |        How often do you have F. salad and vegetables?                                                                                                                                        
18.G  |freq_juices|       How often do you have G. fruit juices?                                                                                                                                            
18.H  | freq_wholem|       How often do you have H. wholemeal products (such as porridge, bread, muesli)?                                                                                                       
19.A  |freq_milk  |       How often do you have A. milk or other dairy products?                                                                                                                                
19.B  |freq_cof|          How often do you have B. coffee?                                                                                                                                                      
19.C  |freq_tea |         How often do you have C. tea?                                                                                                                                                         
19.D  |freq_beer|         How often do you have D. beer?                                                                                                                                                        
19.E  |freq_wine|         How often do you have E. wine?                                                                                                                                                        
19.F  |freq_spirits |     How often do you have F. spirits?                                                                                                                                                        
20    |smoking    |       Do you smoke?                                                                                                                                                                           
21    |phys_cond|         How would you rate your current physical condition?                                                                                                                                     
22    |day_act|           Which of the following statements best describes your typical day (work, studies, leisure)?                                                                                             
23    |h_act_light|       On average, how much time do you spend each day doing light everyday physical activity, such as cycling, walking or rollerblading? Tick the alternative that best describes the average time…
24  |  freq_sports  |     How often do you exercise/do sports in your free time?                                                                                                                                  
26  |  h_act_none   |     On average, how long do you spend each day in your free time on activities that are sedentary (such as doing handicrafts, reading, watching television, playing computer games, surfing the …
S1  |  life_qual  |       How would you rate your quality of life?                                                                                                                                                
S2  |  education  |       What is the highest level of education you have attained?                                                                                                                               
S3  |  employment |       Employment status  


#Remove end point variable

In the inital data analysis stage we do not want to look at our primary endpoint question 5 "How would you rate your recent health in general?" i.e. "health" to avoid fishing for correlations.
```{r}
temp <- data %>% dplyr::select(donor,health) %>% filter(is.na(health))
data <- data %>% filter(!donor %in% unique(temp$donor))  %>% dplyr::select(-health)
data %>% group_by(questionnaire) %>% count()
```

### Remove donors with extreme physiological measures

As decided previously, we remove data according the following criteria:

* CRP >= 30
* Ferritin >= 400 

```{r}
#Rename because next file will also bring an object called data
qdata <-data
#load measurement data
load(file = "../data/quest_2_donor_data.rdata")

data <- data %>% 
  rename(Ferritin_beginning = Ferritin,
         CRP_beginning = CRP) %>% dplyr::select(donor,Ferritin_beginning,CRP_beginning) %>% inner_join(qdata)
cat(nrow(data)/2,"\n")

data <- data %>%   filter(Ferritin_beginning < 400)
cat(nrow(data)/2,"after high ferritin removal\n")
data <- data %>%   filter(CRP_beginning < 30)
cat(nrow(data)/2,"after high CRP removal\n")
data <- data %>% ungroup()
```

```{r}
data <- data %>% dplyr::select(-Ferritin_beginning,-CRP_beginning)
data %>% group_by(questionnaire) %>% count()
```

#Preprocessing recodings
Questions with 3 possible answers : "no", "yes", "don't know", are recoded as an ordered factor with 3 levels: "yes"< "don't know" <  "no".  
```{r}
#Have you felt that you were able to accomplish less than you would have liked to?
temp <- as.character(data$accomplish)
temp[is.na(temp)] <- "don't know"
data$accomplish = ordered(temp,
                levels =  c("yes","don't know", "no"))
#During the past four weeks, has your mental health interfered with your work or other daily activities?
##Later in explorative_analysis.Rmd we find out that some variables cannot be fitted. We remove them now not to loose data over them
# temp <- as.character(data$ment_inf)
# temp[is.na(temp)] <- "don't know"
# data$ment_inf = ordered(temp,
#                 levels =  c("yes","don't know", "no"))

#During the past four weeks, have your physical symptoms (such as pain) or health in general interfered with your work or other daily activities?
temp <- as.character(data$phys_inf)
temp[is.na(temp)] <- "don't know"
data$phys_inf = ordered(temp,
                levels =  c("yes","don't know", "no"))



```



# How much data do we have
```{r}
#Save questions only asked in the second questionaire to be added back later.
additionalq <- data %>% filter(questionnaire == "Second") %>% dplyr::select(donor, life_qual, education, employment)
#drop out variables that, 
# we know from our previous work are too messy: iron_supp_detailed
# were only asked in second questionaire: life_qual, education, employment
# are conveniency variables: AgeGroup, SixMonthsFromStartCount_FB, TwoYearsFromStartCount_FB
temp <- data %>% dplyr::select(-iron_supp_detailed, -life_qual, -education, -employment,-AgeGroup, -SixMonthsFromStartCount_FB,-TwoYearsFromStartCount_FB)

#

temp <- na.omit(temp)
table(table(temp$donor))
```

`r as.vector(table(duplicated(temp$donor))[2])` donors with complete data in both questionaires.


```{r}
temp %>% group_by(questionnaire) %>%  count()
```


As we will loose almost half of our data if we study only complete data we have try to make decisions on which questions to exclude prior to analyses or which data to impute.

```{r}
temp <- data %>% dplyr::select(-iron_supp_detailed, -life_qual, -education, -employment,-AgeGroup, -SixMonthsFromStartCount_FB,-TwoYearsFromStartCount_FB,-donor)
temp <- temp %>% group_by(questionnaire) %>%  summarise_all(funs(sum(is.na(.))))
temp <- as_tibble(t(temp), rownames = "Variable") %>% filter(Variable != "questionnaire")
colnames(temp)[2] <- "First"
colnames(temp)[3] <- "Second"
temp <- temp %>%  gather(key="questionnaire",value="NAs",-Variable) %>% mutate(NAs = as.numeric(gsub(" ","",NAs))) %>% filter(NAs != 0)

ggplot(temp) + geom_col(aes(x=Variable,y=NAs,fill=questionnaire),pos="dodge") +  coord_flip() 

```



```{r}
temp <- temp %>% arrange(NAs) %>% tail(10)

ggplot(temp) + geom_col(aes(x=Variable,y=NAs,fill=questionnaire),pos="dodge") +  coord_flip() 

```

People cannot remember if they have taken vitamine supplements containing iron or vitamine C. Let's first see if some of these will drop out because they correlate heavily with some other explenatory variable.

# Correlations between variables

```{r , warning=FALSE}

temp <- data %>% dplyr::select(-iron_supp_detailed, -life_qual, -education, -employment,-AgeGroup, -SixMonthsFromStartCount_FB,-TwoYearsFromStartCount_FB, -donor,-questionnaire,-sex,-group)
temp$date <- as.numeric(temp$date)

#Change the numeric to factors to enable similar correlation calculation for all
temp <- temp %>% mutate_if(is.numeric, ~cut(.,5))

#https://datascience.stackexchange.com/questions/893/how-to-get-correlation-between-two-categorical-variable-and-a-categorical-variab
cormat <- matrix(NA,ncol = ncol(temp),nrow=ncol(temp))
rownames(cormat) <-colnames(temp)
colnames(cormat) <-colnames(temp)
for (i in rownames(cormat)) {
  for (j in colnames(cormat)) {
    d <- table(as.factor(temp[[i]]),as.factor(temp[[j]]))
#    print(d)
    #drop any all zero rows
    d <- d[apply(X=d,MARGIN = 1,FUN = function(x){!all(x == 0 )}),]
    d <- d[,apply(X=d,MARGIN = 2,FUN = function(x){!all(x == 0 )})]
    #print(d)
    t <- chisq.test(d,correct = F)
    cramersV <- sqrt(t$statistic / sum(d))
    if (cramersV > 1) {
      #sometimes it fails in perfect correlation
      cramersV <- 1
    }
    cormat[i,j] <- cramersV
  }
}


```

Visualise correlation disribution
```{r}
temp <- data.frame(correlation=cormat[lower.tri(cormat)])
ggplot(temp) +  geom_histogram(aes(x=correlation),binwidth = 0.1)

#and create a fake significance matrix of arbitary correlation of 0.5
p.mat <- matrix(1, nrow=nrow(cormat),ncol=ncol(cormat))
p.mat[as.vector(cormat > 0.5)] <- 0.01 
diag(p.mat) <- 1

```


```{r}
corrplot(cormat,type="lower",order="hclust",p.mat=p.mat,sig.level = 0.05,insig = "label_sig",pch.col = 'red',pch.cex = 1)
```

Extract those with correlation above 0.5.
```{r}
#take anything with at least 0.5 correlation
any_above <- function(x){any(x > 0.5) }
temp <-cormat
#empty the diagonal
diag(temp) <- 0
#filter fake significance matrix
temp2 <- p.mat[apply(temp,1, any_above),apply(temp,2, any_above)]
#filter correlation matrix
temp <- temp[apply(temp,1, any_above),apply(temp,2, any_above)]




corrplot(temp,type="lower",order="hclust",p.mat=temp2,sig.level = 0.05,insig = "label_sig",pch.col = 'red',pch.cex = 1)
```

#Choose what to keep
Between those that correlate above 0.5 choose which variable to keep to present the correlaiting variables in the exploratory analysis.

## act_light, act_heavy
```{r}
ggplot(data) + geom_bar(aes(x=act_light))
```

```{r}
ggplot(data) + geom_bar(aes(x=act_heavy))
```

```{r}
table(is.na(data$act_heavy))
table(is.na(data$act_light))
```
act_heavy looses more data, but has much more spread pontentially contaning more information.

Keep "act_heavy". 

##health_c vs phys_cond

```{r}
ggplot(data) + geom_bar(aes(x=health_c))
```
```{r}
ggplot(data) + geom_bar(aes(x=phys_cond))
#Notice that this has been already set from worst to best unlike in the original questionaire
```


```{r}
table(is.na(data$health_c))
table(is.na(data$phys_cond))
```


phys_cond looks better and looses less data. 

##vita_c vs vita_multi

```{r}
ggplot(data) + geom_bar(aes(x=vita_C))
```

```{r}
ggplot(data) + geom_bar(aes(x=vita_multi))
```

```{r}
table(is.na(data$vita_C))
table(is.na(data$vita_multi))
```


vita_multi looses much less data and looks better so vita_C can be dropped.

##sleep_night vs sleep_24h
```{r , warning=FALSE}
ggplot(data) + geom_histogram(aes(x=sleep_night),binwidth = 1)
```
```{r , warning=FALSE}
ggplot(data) + geom_histogram(aes(x=sleep_24h),binwidth = 1)
```

```{r}
table(is.na(data$sleep_night))
table(is.na(data$sleep_24h))
```

Distributions look quite similar, but sleep_night looses less data. 

##freq_fruits vs freq_vege
```{r , warning=FALSE}
ggplot(data) + geom_bar(aes(x=freq_fruits))
```
```{r , warning=FALSE}
ggplot(data) + geom_bar(aes(x=freq_vege))
```

```{r}
table(is.na(data$freq_fruits))
table(is.na(data$freq_vege))
```

freq_fruits look better and looses less data.

##freq_sports vs phys_cond
```{r , warning=FALSE}
ggplot(data) + geom_bar(aes(x=freq_sports))
```
```{r , warning=FALSE}
ggplot(data) + geom_bar(aes(x=phys_cond))
```

```{r}
table(is.na(data$freq_sports))
table(is.na(data$phys_cond))
```

Hard to say anything about the distributions, but physical condition looses less data.

##freq_spirits, freq_beer, freq_wine
```{r , warning=FALSE}
ggplot(data) + geom_bar(aes(x=freq_spirits))
```
```{r , warning=FALSE}
ggplot(data) + geom_bar(aes(x=freq_beer))
```
```{r , warning=FALSE}
ggplot(data) + geom_bar(aes(x=freq_wine))
```
```{r}
table(is.na(data$freq_spirits))
table(is.na(data$freq_beer))
table(is.na(data$freq_wine))
```
Hard to say, but freq_beer looses least data. But this does not cover at all persons that drink mostly cider which could be in particular found in pre-menopausal women.

To have have some kind of proxy over these incomplete but overlapping questions we will create a new variable freq_alc.

```{r}
#Combine the alcohol questions
data$freq_alc <- as.ordered(apply(cbind(data$freq_beer,data$freq_wine,data$freq_spirits),1,max,na.rm=TRUE))
#Make it in to a factor
levels(data$freq_alc) <- levels(data$freq_beer)

```


##freq_cutlets, freq_meat, diet, freq_meat, freq_fish, freq_eggs
```{r , warning=FALSE}
ggplot(data) + geom_bar(aes(x=freq_cutlets))
```
```{r , warning=FALSE}
ggplot(data) + geom_bar(aes(x=freq_red_meat))
```
```{r , warning=FALSE}
ggplot(data) + geom_bar(aes(x=diet))
```
```{r , warning=FALSE}
ggplot(data) + geom_bar(aes(x=freq_fish))
```
```{r , warning=FALSE}
ggplot(data) + geom_bar(aes(x=freq_eggs))
```

```{r}
table(is.na(data$freq_cutlets))
table(is.na(data$freq_red_meat))
table(is.na(data$diet))
table(is.na(data$freq_fish))
table(is.na(data$freq_eggs))
```

As already seen in the enrollment data paper freq_red_meats distribution looks best also it looses leas data.


##BMI, height and weigth
```{r , warning=FALSE}
ggplot(data) + geom_histogram(aes(x=BMI),binwidth = 1)
```
```{r , warning=FALSE}
ggplot(data) + geom_histogram(aes(x=height),binwidth = 5)
```

```{r , warning=FALSE}
ggplot(data) + geom_histogram(aes(x=weight),binwidth = 5)
```

```{r}
table(is.na(data$BMI))
table(is.na(data$height))
table(is.na(data$weight))
```

Keep BMI as composite of both.

##freq_energy, freq_sad, health_soc, freq_calm, accomplish
```{r , warning=FALSE}
ggplot(data) + geom_bar(aes(x=freq_energy))
```
```{r , warning=FALSE}
ggplot(data) + geom_bar(aes(x=freq_sad))
```
```{r , warning=FALSE}
ggplot(data) + geom_bar(aes(x=health_soc))
```
```{r , warning=FALSE}
ggplot(data) + geom_bar(aes(x=freq_calm))
```
```{r , warning=FALSE}
ggplot(data) + geom_bar(aes(x=accomplish))
```

```{r}
table(is.na(data$freq_energy))
table(is.na(data$freq_sad))
table(is.na(data$health_soc))
table(is.na(data$freq_calm))
table(is.na(data$accomplish))
```

accomplish has a tricky distribution.

freq_energy correlates with phys_cond and can be dropped for that.

health_soc has a very skewed distiribution.

freq_calm looses least data and has maybe the best looking distribution.

##h_act_light vs h_act_noen
```{r , warning=FALSE}
ggplot(data) + geom_bar(aes(x=h_act_light))
```
```{r , warning=FALSE}
ggplot(data) + geom_bar(aes(x=h_act_none))
```

```{r}
table(is.na(data$h_act_light))
table(is.na(data$h_act_none))
```

h_act_light looses less data.

#Drop questions based on correlation analysis
Based on the above we have decided to drop: act_light, health_c, vita_C, sleep_24h, freq_vege, freq_sports, freq_spirits, freq_beer, freq_wine, freq_cutlets, diet, freq_fish, freq_eggs, weight, height, accomplish, freq_energy, health_soc, freq_sad, h_act_none

```{r}
temp <- data %>% dplyr::select(-iron_supp_detailed, -life_qual, -education, -employment,-AgeGroup, -SixMonthsFromStartCount_FB,-TwoYearsFromStartCount_FB, -act_light, -health_c, -vita_C, -sleep_24h, -freq_vege, -freq_sports, -freq_spirits, -freq_wine, -freq_beer, -freq_cutlets, -diet, -freq_fish, -freq_eggs, -weight, -height, -accomplish, -freq_energy, -health_soc, -freq_sad,-h_act_none)

#Later in explorative_analysis.Rmd we find out that some variables cannot be fitted. We remove them now not to loose data over them
temp <- temp %>% dplyr::select(-sleep_night,
-ment_inf,
-vita_iron
#education is removed later as it was only asked in the second questionaire
#smoking_behavior # this is brought in and removed in
)

temp <- na.omit(temp)
temp %>% group_by(questionnaire) %>% count()
```
From `r ncol(temp)` questions.

```{r}

#But questionnaire, donor, group, age, sex and date are not procesed into change between questionnaires
tx <- setdiff(colnames(temp),c("questionnaire", "donor", "group", "age", "sex","date" ))
tx
```


```{r}
length(tx)
```

How will the sex/menopause groups look
```{r}
temp %>% group_by(questionnaire,group) %>% filter(group != 'Women_pre_menop_no_mens') %>%  count()
```

```{r}
table(table(temp$donor))
```

`r as.vector(table(duplicated(temp$donor))[2])` donors with complete data in both questionaires.


By excluding several questions that correlate with other questions and have distributions that are skewed (e.g. diet, accomplish) or are uniformative (e.g. act_light)  or are not clearly ordered factors (like diet) we get complete data for about two hundred more persons. There might still be too many explenatory variables for the smallest group i.e. Post_menopause_women and we will still loose data when we look for complete cases in both questionaires.

#Recode answers as change between the questionnaires

Exclude questions that are dealt with later in the regression data processing:
BMI, smoking health, date, age, group. "health" has been already excluded. N.B. smoking is not part of the hypothesis testing. It is just processed later.
```{r}
temp <- temp %>% dplyr::select(-BMI,-smoking,-date,-age,-group,-sex)
#Extract complete data for both questinaires and separate questionaire answers into different columns for processing
splitbyq <- split(temp,f=temp$questionnaire)
first <- splitbyq[["First"]]
second <- splitbyq[["Second"]]
colnames(second) <- paste0(colnames(second),"_2")
temp <- inner_join(first,second,by=c("donor"="donor_2"))
cat(nrow(temp),"complete cases in both questionnaires\n")
```



```{r}

classes <- sort(unlist(lapply(lapply(first,class),paste,collapse=" ")))
classes <- as.tibble(cbind(names(classes),classes))
no_answers <- first %>% summarise_all(~length(levels(.)))
no_answers <- as_tibble(cbind(variable = names(no_answers), t(no_answers)))
no_answers <- full_join(no_answers,classes,by=c("variable"="V1")) %>% rename(no_answers=V2)
no_answers %>% group_by(no_answers)  %>% count() %>%  arrange(n)
```

```{r}
no_answers %>%  filter(no_answers == 0)
```

no_aswers == 0 are non-factoral variables. Change in cholesterol, blood_sugar, HBP, low_hb_hist, anemia_hist, high_hb_hist is coded as if the status changed during the study.

```{r}
temp <- temp %>% mutate(cholesterol_d = case_when(cholesterol == 0 & cholesterol_2 == 1 ~ 1,
                                          TRUE ~ 0
                                          )) %>% mutate(cholesterol_d = as.logical(cholesterol_d))

temp <- temp %>% mutate(blood_sugar_d = case_when(blood_sugar == 0 & blood_sugar_2 == 1 ~ 1,
                                          TRUE ~ 0
                                          )) %>% mutate(blood_sugar_d = as.logical(blood_sugar_d))

temp <- temp %>% mutate(HBP_d = case_when(HBP == 0 & HBP_2 == 1 ~ 1,
                                          TRUE ~ 0
                                          )) %>% mutate(HBP_d = as.logical(HBP_d))

temp <- temp %>% mutate(low_hb_hist_d = case_when(low_hb_hist == 0 & low_hb_hist_2 == 1 ~ 1,
                                          TRUE ~ 0
                                          )) %>% mutate(low_hb_hist_d = as.logical(low_hb_hist_d))

temp <- temp %>% mutate(anemia_hist_d = case_when(anemia_hist == 0 & anemia_hist_2 == 1 ~ 1,
                                          TRUE ~ 0
                                          )) %>% mutate(anemia_hist_d = as.logical(anemia_hist_d))

temp <- temp %>% mutate(high_hb_hist_d = case_when(high_hb_hist == 0 & high_hb_hist_2 == 1 ~ 1,
                                          TRUE ~ 0
                                          )) %>% mutate(high_hb_hist_d = as.logical(high_hb_hist_d))
```


iron_supp recoded as either reducing, stable or increasing.

```{r}
temp <- temp %>% mutate(iron_supp_d = case_when(iron_supp == FALSE & iron_supp_2 == TRUE ~ "Increased",
                                                iron_supp == TRUE & iron_supp_2 == FALSE ~ "Reduced",
                                          TRUE ~ "Stable"
                                          )) %>% mutate( iron_supp_d= factor(iron_supp_d,levels = c("Reduced","Stable","Increased"),ordered = TRUE))

```

sleep_night recoded as difference

```{r}
##Later in explorative_analysis.Rmd we find out that some variables cannot be fitted. We remove them now not to loose data over them
#temp <- temp %>% mutate(sleep_night_d = sleep_night_2 - sleep_night)
```

```{r}
no_answers %>%  filter(no_answers == 6)
```

These are all ordered factors. They are recoded as  either reducing, stable or increasing.

```{r}
temp <- temp %>% mutate(freq_calm_d = as.numeric(freq_calm_2)  - as.numeric(freq_calm)) %>% 
  mutate(freq_calm_d = case_when(
    freq_calm_d > 0 ~ "Increased",
    freq_calm_d < 0 ~ "Reduced",
    TRUE ~ "Stable"
  )) %>% mutate( freq_calm_d= factor(freq_calm_d,levels = c("Reduced","Stable","Increased"),ordered = TRUE))

temp <- temp %>% mutate(freq_red_meat_d = as.numeric(freq_red_meat_2)  - as.numeric(freq_red_meat)) %>% 
  mutate(freq_red_meat_d = case_when(
    freq_red_meat_d > 0 ~ "Increased",
    freq_red_meat_d < 0 ~ "Reduced",
    TRUE ~ "Stable"
  )) %>% mutate( freq_red_meat_d= factor(freq_red_meat_d,levels = c("Reduced","Stable","Increased"),ordered = TRUE))

temp <- temp %>% mutate(freq_fruits_d = as.numeric(freq_fruits_2)  - as.numeric(freq_fruits)) %>% 
  mutate(freq_fruits_d = case_when(
    freq_fruits_d > 0 ~ "Increased",
    freq_fruits_d < 0 ~ "Reduced",
    TRUE ~ "Stable"
  )) %>% mutate( freq_fruits_d= factor(freq_fruits_d,levels = c("Reduced","Stable","Increased"),ordered = TRUE))

temp <- temp %>% mutate(freq_juices_d = as.numeric(freq_juices_2)  - as.numeric(freq_juices)) %>% 
  mutate(freq_juices_d = case_when(
    freq_juices_d > 0 ~ "Increased",
    freq_juices_d < 0 ~ "Reduced",
    TRUE ~ "Stable"
  )) %>% mutate( freq_juices_d= factor(freq_juices_d,levels = c("Reduced","Stable","Increased"),ordered = TRUE))

temp <- temp %>% mutate(freq_wholem_d = as.numeric(freq_wholem_2)  - as.numeric(freq_wholem)) %>% 
  mutate(freq_wholem_d = case_when(
    freq_wholem_d > 0 ~ "Increased",
    freq_wholem_d < 0 ~ "Reduced",
    TRUE ~ "Stable"
  )) %>% mutate( freq_wholem_d= factor(freq_wholem_d,levels = c("Reduced","Stable","Increased"),ordered = TRUE))

temp <- temp %>% mutate(freq_milk_d = as.numeric(freq_milk_2)  - as.numeric(freq_milk)) %>% 
  mutate(freq_milk_d = case_when(
    freq_milk_d > 0 ~ "Increased",
    freq_milk_d < 0 ~ "Reduced",
    TRUE ~ "Stable"
  )) %>% mutate( freq_milk_d= factor(freq_milk_d,levels = c("Reduced","Stable","Increased"),ordered = TRUE))

temp <- temp %>% mutate(freq_cof_d = as.numeric(freq_cof_2)  - as.numeric(freq_cof)) %>% 
  mutate(freq_cof_d = case_when(
    freq_cof_d > 0 ~ "Increased",
    freq_cof_d < 0 ~ "Reduced",
    TRUE ~ "Stable"
  )) %>% mutate( freq_cof_d= factor(freq_cof_d,levels = c("Reduced","Stable","Increased"),ordered = TRUE))

temp <- temp %>% mutate(freq_tea_d = as.numeric(freq_tea_2)  - as.numeric(freq_tea)) %>% 
  mutate(freq_tea_d = case_when(
    freq_tea_d > 0 ~ "Increased",
    freq_tea_d < 0 ~ "Reduced",
    TRUE ~ "Stable"
  )) %>% mutate( freq_tea_d= factor(freq_tea_d,levels = c("Reduced","Stable","Increased"),ordered = TRUE))

temp <- temp %>% mutate(phys_cond_d = as.numeric(phys_cond_2)  - as.numeric(phys_cond)) %>% 
  mutate(phys_cond_d = case_when(
    phys_cond_d > 0 ~ "Increased",
    phys_cond_d < 0 ~ "Reduced",
    TRUE ~ "Stable"
  )) %>% mutate( phys_cond_d= factor(phys_cond_d,levels = c("Reduced","Stable","Increased"),ordered = TRUE))

```




```{r}
no_answers %>%  filter(no_answers == 4)
```


These are all ordered factors. They are recoded as  either reducing, stable or increasing.
```{r}
temp <- temp %>% mutate(act_heavy_d = as.numeric(act_heavy_2)  - as.numeric(act_heavy)) %>% 
  mutate(act_heavy_d = case_when(
    act_heavy_d > 0 ~ "Increased",
    act_heavy_d < 0 ~ "Reduced",
    TRUE ~ "Stable"
  )) %>% mutate( act_heavy_d= factor(act_heavy_d,levels = c("Reduced","Stable","Increased"),ordered = TRUE))

#sleep_enough
temp <- temp %>% mutate(sleep_enough_d = as.numeric(sleep_enough_2)  - as.numeric(sleep_enough)) %>% 
  mutate(sleep_enough_d = case_when(
    sleep_enough_d > 0 ~ "Increased",
    sleep_enough_d < 0 ~ "Reduced",
    TRUE ~ "Stable"
  )) %>% mutate( sleep_enough_d= factor(sleep_enough_d,levels = c("Reduced","Stable","Increased"),ordered = TRUE))

#day_act
temp <- temp %>% mutate(day_act_d = as.numeric(day_act_2)  - as.numeric(day_act)) %>% 
  mutate(day_act_d = case_when(
    day_act_d > 0 ~ "Increased",
    day_act_d < 0 ~ "Reduced",
    TRUE ~ "Stable"
  )) %>% mutate( day_act_d= factor(day_act_d,levels = c("Reduced","Stable","Increased"),ordered = TRUE))

#h_act_light
temp <- temp %>% mutate(h_act_light_d = as.numeric(h_act_light_2)  - as.numeric(h_act_light)) %>% 
  mutate(h_act_light_d = case_when(
    h_act_light_d > 0 ~ "Increased",
    h_act_light_d < 0 ~ "Reduced",
    TRUE ~ "Stable"
  )) %>% mutate( h_act_light_d= factor(h_act_light_d,levels = c("Reduced","Stable","Increased"),ordered = TRUE))

#vita_multi
temp <- temp %>% mutate(vita_multi_d = as.numeric(vita_multi_2)  - as.numeric(vita_multi)) %>% 
  mutate(vita_multi_d = case_when(
    vita_multi_d > 0 ~ "Increased",
    vita_multi_d < 0 ~ "Reduced",
    TRUE ~ "Stable"
  )) %>% mutate( vita_multi_d= factor(vita_multi_d,levels = c("Reduced","Stable","Increased"),ordered = TRUE))

# Later in explorative_analysis.Rmd we find out that some variables cannot be fitted. We remove them now not to loose data over them
#vita_iron
# temp <- temp %>% mutate(vita_iron_d = as.numeric(vita_iron_2)  - as.numeric(vita_iron)) %>% 
#   mutate(vita_iron_d = case_when(
#     vita_iron_d > 0 ~ "Increased",
#     vita_iron_d < 0 ~ "Reduced",
#     TRUE ~ "Stable"
#   )) %>% mutate( vita_iron_d= factor(vita_iron_d,levels = c("Reduced","Stable","Increased"),ordered = TRUE))

```

But these 2 are diffent order of time and they are turned around to have same reducing, stable or increasing coding.

sleep_tired Levels: always_or_almost < mostly < rarely < never i.e. is their tiredness reduced, stable or increasing

anti_inf: Levels: daily < not_daily < occasionally < none

```{r}
temp <- temp %>% mutate(sleep_tired_d = as.numeric(sleep_tired)  - as.numeric(sleep_tired_2)) %>% 
  mutate(sleep_tired_d = case_when(
    sleep_tired_d > 0 ~ "Increased",
    sleep_tired_d < 0 ~ "Reduced",
    TRUE ~ "Stable"
  )) %>% mutate( sleep_tired_d= factor(sleep_tired_d,levels = c("Reduced","Stable","Increased"),ordered = TRUE))

temp <- temp %>% mutate(anti_inf_d = as.numeric(anti_inf)  - as.numeric(anti_inf_2)) %>% 
  mutate(anti_inf_d = case_when(
    anti_inf_d > 0 ~ "Increased",
    anti_inf_d < 0 ~ "Reduced",
    TRUE ~ "Stable"
  )) %>% mutate( anti_inf_d= factor(anti_inf_d,levels = c("Reduced","Stable","Increased"),ordered = TRUE))


```


```{r}
no_answers %>%  filter(no_answers == 3)
```

phys_inf Levels: yes < don't know < no -> has your physical health inferior? e.g. first : no, second: yes -> physical health inferring "Increased"

ment_inf Levels: yes < don't know < no -> has your mental health inferred?

```{r}
temp <- temp %>% mutate(phys_inf_d = as.numeric(phys_inf)  - as.numeric(phys_inf_2)) %>% 
  mutate(phys_inf_d = case_when(
    phys_inf_d > 0 ~ "Increased",
    phys_inf_d < 0 ~ "Reduced",
    TRUE ~ "Stable"
  )) %>% mutate( phys_inf_d= factor(phys_inf_d,levels = c("Reduced","Stable","Increased"),ordered = TRUE))

#Later in explorative_analysis.Rmd we find out that some variables cannot be fitted. We remove them now not to loose data over them
# temp <- temp %>% mutate(ment_inf_d = as.numeric(ment_inf)  - as.numeric(ment_inf_2)) %>% 
#   mutate(ment_inf_d = case_when(
#     ment_inf_d > 0 ~ "Increased",
#     ment_inf_d < 0 ~ "Reduced",
#     TRUE ~ "Stable"
#   )) %>% mutate( ment_inf_d= factor(ment_inf_d,levels = c("Reduced","Stable","Increased"),ordered = TRUE))


```



```{r}
no_answers %>%  filter(no_answers == 5)
```

freq_alc Levels: daily_or_almost < a_few_per_week < a_few_per_month < very_rarely < never 

```{r}
temp <- temp %>% mutate(freq_alc_d = as.numeric(freq_alc)  - as.numeric(freq_alc_2)) %>% 
  mutate(freq_alc_d = case_when(
    freq_alc_d > 0 ~ "Increased",
    freq_alc_d < 0 ~ "Reduced",
    TRUE ~ "Stable"
  )) %>% mutate( freq_alc_d= factor(freq_alc_d,levels = c("Reduced","Stable","Increased"),ordered = TRUE))

```

Select only the change variables and store.

```{r}
cols2keep<- grep("_d$",colnames(temp),value = TRUE)
cols2keep
```
```{r}
cols2keep<- c("donor",cols2keep)
changedata <- temp %>% dplyr::select(cols2keep)
dim(changedata)
```

#Checks
```{r}
setdiff(colnames(first),gsub("_d$","",colnames(changedata)))
```

```{r}
setdiff(gsub("_d$","",colnames(changedata)),colnames(first))
```


```{r}
classes <- sort(unlist(lapply(lapply(changedata,class),paste,collapse=" ")))
classes <- as.tibble(cbind(names(classes),classes))
levels <- unlist(lapply(lapply(changedata,levels),paste,collapse=" "))
levels <- as.tibble(cbind(names(levels),levels))
changedataclasses <- full_join(classes,levels,by=c("V1"="V1")) %>% rename(variable=V1)  
kable(changedataclasses,caption="Classes and levels of variables only analysed in explorative analysis")
```

#Additional questions of the second questionaire


Consider adding the second questionaire additional questions back and save
```{r}
additionalq <- inner_join(changedata,additionalq,by=c("donor"="donor"))
table(is.na(additionalq$life_qual))
table(is.na(additionalq$employment))
table(is.na(additionalq$education))

```

```{r}
ggplot(additionalq) + geom_bar(aes(x=life_qual))
```

Could excellent be combined very_good and very_poor and poor to average?
```{r}
additionalq <- additionalq %>% mutate(life_qual = case_when(
  life_qual == "very_poor" ~ "average",
  life_qual == "poor" ~ "average",
  life_qual == "average" ~ "average",
  life_qual == "good" ~ "good",
  life_qual == "very_good" ~ "very_good",
  life_qual == "excellent" ~ "very_good"
)) %>% mutate(life_qual = factor(life_qual))
```


```{r}
ggplot(additionalq) + geom_bar(aes(x=life_qual))
```


```{r}
ggplot(additionalq) + geom_bar(aes(x=employment))
```

This is not at all informative. Could it be fulltime or not?

```{r}
additionalq <- additionalq %>% mutate(employment = case_when(
  employment == "full-time" ~ TRUE,
  TRUE ~ FALSE
  ))
```


```{r}
ggplot(additionalq) + geom_bar(aes(x=education))
```

Also some grouping would be good in here.

```{r}
additionalq <- additionalq %>% mutate(education = case_when(
  education == "trade-school" |  education == "vocational_college" |  education == "graduate_degree" | education == "bachelor_degree"   ~ "high",
  education == "vocational_school" |  education == "high-school"  ~ "middle",
  education == "basic_education" | education == "middle-school"  ~ "lower"
  
  )) %>% mutate(education = factor(education))
#Later in explorative_analysis.Rmd we find out that education cannot be fitted 
additionalq <- additionalq %>% select(-education)
```


```{r}

cat("Down to:" ,nrow(na.omit(additionalq)),"from:",nrow(additionalq),"\n")
changedata <- na.omit(additionalq)


#changedata <- inner_join(changedata,additionalq,by=c("donor"="donor"))
file <- "../results/changedata.rdata"
save(changedata,file=file)

```
Just to check:
```{r}
changedata %>% count()
```

By dropping many questions we manage to save several hundred donors. In addition, next in explorative_analysis.Rmd even couple more variables need to dropped out in order to fit the models. Hence, dropping related variables is required.