---
title: "Selection bias analysis for Associations between changes in self-rated health, lifestyle and iron levels in Finnish blood donors"
author: "Mikko Arvas"
date: "`r Sys.time()`"
output: html_document
---

```{r setup, include=FALSE}
Sys.setlocale("LC_MESSAGES", 'en_GB.UTF-8')
Sys.setenv(LANG = "en_US.UTF-8")
# echo "rmarkdown::render('selection_bias.Rmd', clean=TRUE,output_format='html_document',output_file='../results/selection_bias.html')" | R --slave
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
library(tidyverse)
library(knitr)
```



# Executive summary

We collect participants of different analysis subcohorts and plot change in their health status between the first and second health questionaires. Furthermore, for each inital health class we count the rate of participation for the two analysis subcohorts. 

# Load the data

Load enrollment data
```{r}
file <- "../data/first_questionnaire_data_.rdata"
load(file)
count(first_questionnaire_data)
```

And 4 persons that did not answer the first questionnaire. 

Load combined data

```{r loading}
#Load questionnaire data
load(file = "../data/final_data.rdata") 
data <- data %>% ungroup()
data %>% group_by(questionnaire) %>% count()

```

```{r}
data %>% group_by(questionnaire,group)  %>%  count()
```

Load the explorative analysis data

```{r}
#Load questionnaire data recoded to change
load(file = "../results/changedata.rdata") #  changedata 


```


# Make exclusions as in hypothesis testing and explorative analysis

```{r}

file <- "../results/hypregdata_unscaled.rdata"
load(file)

hyp_health <- data %>% inner_join(regression_data %>% dplyr::select(donor))

hyp_health <- hyp_health %>%
  dplyr::select(donor, sex, health, questionnaire) %>% 
  spread(key = questionnaire, value = health) 
  #this creates NA's from nowhere?
  #%>% filter(First != "NA" & Second != "NA") 
  
hyp_health %>%  count(sex)
```


```{r}
hyp_health  %>% count()
```

```{r}

hyp_health_sum <- hyp_health %>% group_by(sex) %>% count(First,Second) %>%  mutate(pros = paste0(round(n / sum(n) *100,0), " %"),
                                                                                     freq = n / sum(n))
hyp_health_sum  %>% select(sex,First,Second,pros)
```


```{r}
expl_health <- hyp_health %>%  inner_join( changedata %>% select(donor))
expl_health %>% count()
```

```{r}
expl_health %>% count(sex)
```

```{r}

expl_health_sum <- expl_health %>% group_by(sex) %>% count(First,Second) %>%  mutate(pros = paste0(round(n / sum(n) *100,0), " %"),
                                                                                     freq = n / sum(n))
expl_health_sum %>% select(sex,First,Second,pros)

```

# Plot

```{r fig.height=6, fig.width=10, warning=FALSE}
p <- hyp_health %>% 
    mutate(evolution = case_when(First < Second ~"improved",
                               First == Second ~ "same",
                               TRUE ~ "worsened")) %>% 
  ggplot(aes (x = First, y = Second, color = evolution)) +
  geom_count() +
  facet_grid(~sex) +
  geom_text(data = hyp_health_sum, 
              aes(x = First, y = Second, label = pros), color = "black") +
      #https://stackoverflow.com/questions/55531147/how-to-add-labels-to-facet-wraped-geom-count-plot
   #you have to make the counts first if you want to overlay data and counts
  scale_color_manual(values = c("lightgreen", "gray", "pink"),name="Evolution") +
  theme_bw() +
  xlab("First questionnaire") +
  ylab("Second questionnaire")

 p
 filename <- "../results/figures/health_group_change_1416.pdf"
 ggsave(filename=filename, p, width = 180,  height = 90,units="mm", dpi=600, scale=1.0)
 
```


```{r fig.height=6, fig.width=10, warning=FALSE}
p <- expl_health %>% 
    mutate(evolution = case_when(First < Second ~"improved",
                               First == Second ~ "same",
                               TRUE ~ "worsened")) %>% 
  ggplot(aes (x = First, y = Second, color = evolution)) +
  geom_count() +
  facet_grid(~sex) +
  geom_text(data = hyp_health_sum, 
              aes(x = First, y = Second, label = pros), color = "black") +
        scale_color_manual(values = c("lightgreen", "gray", "pink"),name="Evolution") +
  theme_bw() +
  xlab("First questionnaire") +
  ylab("Second questionnaire")

 p
 filename <- "../results/figures/health_group_change_1045.pdf"
 ggsave(filename=filename, p, width = 180,  height = 90,units="mm", dpi=600, scale=1.0)
 
```

Let's put them together
```{r}
hyp_health_sum <-  hyp_health_sum %>% mutate(Group = "Hypothesis")
expl_health_sum <- expl_health_sum %>% mutate(Group = "Explorative")
health_sum <- bind_rows(hyp_health_sum, expl_health_sum)

hyp_health <-  hyp_health %>% mutate(Group = "Hypothesis")
expl_health <- expl_health %>% mutate(Group = "Explorative")
health <- bind_rows(hyp_health, expl_health)


```

```{r fig.height=6, fig.width=10, warning=FALSE}
p <-  health %>%
    mutate(evolution = case_when(First < Second ~"improved",
                               First == Second ~ "same",
                               TRUE ~ "worsened")) %>% 
  ggplot(aes (x = First, y = Second, color = evolution)) +
  geom_count() +
  facet_grid(Group~sex) +
  geom_text(data = health_sum, 
              aes(x = First, y = Second, label = pros), color = "black") +
      #https://stackoverflow.com/questions/55531147/how-to-add-labels-to-facet-wraped-geom-count-plot
   #you have to make the counts first
  scale_color_manual(values = c("lightgreen", "gray", "pink"),name="Evolution") +
  theme_bw() +
  xlab("First questionnaire") +
  ylab("Second questionnaire")

 p
 filename <- "../results/figures/health_group_change.pdf"
 ggsave(filename=filename, p, width = 180,  height = 140,units="mm", dpi=600, scale=1.0)
 
```




# Completion rate table
```{r}
enrol_comp <- first_questionnaire_data %>% rename( Health = QR17, Sex=sex)%>%  group_by(Sex,Health) %>%  summarise(n = n()) %>% mutate(
                                                          freq = n / sum(n),
                                                          pros = paste0(round(n / sum(n) *100,0), " %")
                                                          
                                                          )
enrol_comp  
```

```{r}
hyp_comp <- first_questionnaire_data %>% inner_join(hyp_health %>% select(donor),by = c("donor"="donor")) %>% rename( Health = QR17, Sex=sex)%>%  group_by(Sex,Health) %>%  summarise(n = n()) %>% mutate(
                                                          freq = n / sum(n),
                                                          pros = paste0(round(n / sum(n) *100,0), " %")
                                                          ) %>% select(Health, pros) %>% rename(Hyp_comp = pros)
hyp_comp
```

```{r}
expl_comp <- first_questionnaire_data %>% inner_join(expl_health %>% select(donor),by = c("donor"="donor")) %>% rename( Health = QR17, Sex=sex)%>%  group_by(Sex,Health) %>%  summarise(n = n()) %>% mutate(
                                                          freq = n / sum(n),
                                                          pros = paste0(round(n / sum(n) *100,0), " %")
                                                          ) %>% select(Health,pros) %>% rename(Exp_comp = pros)
expl_comp
```

```{r}
comp <- full_join(enrol_comp,hyp_comp,by=c("Sex"="Sex","Health"="Health")) %>% full_join(expl_comp,by=c("Sex"="Sex","Health"="Health")) %>% select(-freq)
comp
```

Of those who were included in the two subcohorts how many has a specific self-estimated health in the enrollment data.

But how many of those who were included in the two subcohorts ended up in a subcohort.

```{r}
hy <- hyp_health %>% mutate(Hypothesis = "Hypothesis") %>% select(donor,Hypothesis)
ex <- expl_health %>% mutate(Explorative = "Explorative") %>% select(donor,Explorative) 
comp <- first_questionnaire_data %>% rename( Health = QR17, Sex=sex)%>% select(donor,Health,Sex) %>% full_join(hy,by=c("donor"="donor")) %>% full_join(ex,by=c("donor"="donor")) %>% mutate(Subcohort = paste0(Hypothesis,'_',Explorative)) %>% select(-Hypothesis,-Explorative)

comp_sum<- comp %>% group_by(Sex,Health,Subcohort) %>% 
  summarise(n = n()) %>% 
  pivot_wider(names_from = Subcohort, values_from=n) %>% 
  mutate(Hypothesis = Hypothesis_NA + Hypothesis_Explorative,Total=sum(NA_NA ,Hypothesis_NA ,Hypothesis_Explorative,na.rm = TRUE)) %>% 
  rename(Explorative=Hypothesis_Explorative) %>%
  mutate(Hypothesis=replace_na(Hypothesis,0),
         Explorative=replace_na(Explorative,0)
         ) %>% 
  mutate("Hypothesis %"  = paste(round(Hypothesis / Total*100,0), " %"),
         "Explorative %" = paste0(round(Explorative / Total*100,0), " %")
         ) %>% 
#  select(Sex, Health,Total, Hypothesis,"Hypothesis %" ,Explorative,"Explorative %") 
    select(Sex, Health,Total,"Hypothesis %" ,"Explorative %")
comp_sum
```

```{r}
file <- "../results/completion_rates.csv"
write.table(comp_sum,file=file,sep="\t",row.names = FALSE)
```

