---
title: "Selection bias analysis for Associations between changes in self-rated health, lifestyle and iron levels in Finnish blood donors"
author: "Mikko Arvas"
date: "`r Sys.time()`"
output: html_document
---

TODO:
- from dots to to count + percentage symbols
- separate plots for all, reg set, and exp set? everyone to a single facet plot?
- compare counts vectors of all to hyp and exp set

```{r setup, include=FALSE}
Sys.setlocale("LC_MESSAGES", 'en_GB.UTF-8')
Sys.setenv(LANG = "en_US.UTF-8")
# echo "rmarkdown::render('inital_data_analysis.Rmd', clean=TRUE,output_format='html_document',output_file='../results/selection_bias_analysis.html')" | R --slave
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
library(tidyverse)
library(knitr)
```



# Executive summary



# Load the data

Load enrollment data
```{r}
file <- "../data/first_questionnaire_data_.rdata"
load(file)
count(first_questionnaire_data)
```

And 4 persons that did not answer the first questinnaire. 

Load combined data

```{r loading}
#Load questionnaire data
load(file = "../data/final_data.rdata") 
data <- data %>% ungroup()
data %>% group_by(questionnaire) %>% count()

```

```{r}
data %>% group_by(questionnaire,group)  %>%  count()
```

Load the explorative analysis data

```{r}
#Load questionnaire data recoded to change
load(file = "../results/changedata.rdata") #  changedata 


```


# Make exclusions as in hypothesis testing and explorative analysis

```{r}

file <- "../results/hypregdata_unscaled.rdata"
load(file)

hyp_health <- data %>% inner_join(regression_data %>% dplyr::select(donor))

hyp_health <- hyp_health %>%
  dplyr::select(donor, sex, health, questionnaire) %>% 
  spread(key = questionnaire, value = health) 
  #this creates NA's from nowhere?
  #%>% filter(First != "NA" & Second != "NA") 
  
hyp_health %>%  count(sex)
```


```{r}
hyp_health  %>% count()
```

```{r}

hyp_health_sum <- hyp_health %>% group_by(sex) %>% count(First,Second) %>%  mutate(pros = paste0(round(n / sum(n) *100,0), " %"),
                                                                                     freq = n / sum(n))
hyp_health_sum
```


```{r}
expl_health <- hyp_health %>%  inner_join( changedata %>% select(donor))
expl_health %>% count()
```

```{r}
expl_health %>% count(sex)
```

```{r}

expl_health_sum <- expl_health %>% group_by(sex) %>% count(First,Second) %>%  mutate(pros = paste0(round(n / sum(n) *100,0), " %"),
                                                                                     freq = n / sum(n))
expl_health_sum

```

# Plot

```{r fig.height=6, fig.width=10, warning=FALSE}
p <- hyp_health %>% 
    mutate(evolution = case_when(First < Second ~"improved",
                               First == Second ~ "same",
                               TRUE ~ "worsened")) %>% 
  ggplot(aes (x = First, y = Second, color = evolution)) +
  geom_count() +
  facet_grid(~sex) +
  geom_text(data = hyp_health_sum, 
              aes(x = First, y = Second, label = pros), color = "black") +
      #https://stackoverflow.com/questions/55531147/how-to-add-labels-to-facet-wraped-geom-count-plot
   #you have to make the counts first
  scale_color_manual(values = c("lightgreen", "gray", "pink"),name="Evolution") +
  theme_bw() +
  xlab("First questionnaire") +
  ylab("Second questionnaire")

 p
 filename <- "../results/figures/health_group_change_1416.pdf"
 ggsave(filename=filename, p, width = 180,  height = 90,units="mm", dpi=600, scale=1.0)
 
```


```{r fig.height=6, fig.width=10, warning=FALSE}
p <- expl_health %>% 
    mutate(evolution = case_when(First < Second ~"improved",
                               First == Second ~ "same",
                               TRUE ~ "worsened")) %>% 
  ggplot(aes (x = First, y = Second, color = evolution)) +
  geom_count() +
  facet_grid(~sex) +
  geom_text(data = hyp_health_sum, 
              aes(x = First, y = Second, label = pros), color = "black") +
      #https://stackoverflow.com/questions/55531147/how-to-add-labels-to-facet-wraped-geom-count-plot
   #you have to make the counts first
  scale_color_manual(values = c("lightgreen", "gray", "pink"),name="Evolution") +
  theme_bw() +
  xlab("First questionnaire") +
  ylab("Second questionnaire")

 p
 filename <- "../results/figures/health_group_change_1045.pdf"
 ggsave(filename=filename, p, width = 180,  height = 90,units="mm", dpi=600, scale=1.0)
 
```

Let's put them together
```{r}
hyp_health_sum <-  hyp_health_sum %>% mutate(Group = "Hypothesis")
expl_health_sum <- expl_health_sum %>% mutate(Group = "Explorative")
health_sum <- bind_rows(hyp_health_sum, expl_health_sum)

hyp_health <-  hyp_health %>% mutate(Group = "Hypothesis")
expl_health <- expl_health %>% mutate(Group = "Explorative")
health <- bind_rows(hyp_health, expl_health)


```

```{r fig.height=6, fig.width=10, warning=FALSE}
p <-  health %>%
    mutate(evolution = case_when(First < Second ~"improved",
                               First == Second ~ "same",
                               TRUE ~ "worsened")) %>% 
  ggplot(aes (x = First, y = Second, color = evolution)) +
  geom_count() +
  facet_grid(Group~sex) +
  geom_text(data = health_sum, 
              aes(x = First, y = Second, label = pros), color = "black") +
      #https://stackoverflow.com/questions/55531147/how-to-add-labels-to-facet-wraped-geom-count-plot
   #you have to make the counts first
  scale_color_manual(values = c("lightgreen", "gray", "pink"),name="Evolution") +
  theme_bw() +
  xlab("First questionnaire") +
  ylab("Second questionnaire")

 p
 filename <- "../results/figures/health_group_change.pdf"
 ggsave(filename=filename, p, width = 180,  height = 140,units="mm", dpi=600, scale=1.0)
 
```




# Completion rate table
```{r}
enrol_comp <- enrol_health <- first_questionnaire_data %>% rename( Health = QR17, Sex=sex)%>%  group_by(Sex,Health) %>%  summarise(n = n()) %>% mutate(
                                                          freq = n / sum(n),
                                                          pros = paste0(round(n / sum(n) *100,0), " %")
                                                          
                                                          )
enrol_comp  
```

```{r}
hyp_comp <- first_questionnaire_data %>% inner_join(hyp_health %>% select(donor),by = c("donor"="donor")) %>% rename( Health = QR17, Sex=sex)%>%  group_by(Sex,Health) %>%  summarise(n = n()) %>% mutate(
                                                          freq = n / sum(n),
                                                          pros = paste0(round(n / sum(n) *100,0), " %")
                                                          ) %>% select(Health, pros) %>% rename(Hyp_comp = pros)
hyp_comp
```

```{r}
expl_comp <- first_questionnaire_data %>% inner_join(expl_health %>% select(donor),by = c("donor"="donor")) %>% rename( Health = QR17, Sex=sex)%>%  group_by(Sex,Health) %>%  summarise(n = n()) %>% mutate(
                                                          freq = n / sum(n),
                                                          pros = paste0(round(n / sum(n) *100,0), " %")
                                                          ) %>% select(Health,pros) %>% rename(Exp_comp = pros)
expl_comp
```

```{r}
comp <- full_join(enrol_comp,hyp_comp,by=c("Sex"="Sex","Health"="Health")) %>% full_join(expl_comp,by=c("Sex"="Sex","Health"="Health")) %>% select(-freq)
comp
```

Of those who were included in the two subcohorts how many has a specific self-estimated health in the enrollment data.

But how many of those who were included in the two subcohorts ended up in a subcohort.



